### 2014-09-02 ants2015.Rmd
ENH: new stuff talk

---
title: <span style="color:red;">How to build a house with<br />Advanced Normalization Tools (*ANTs*)</span>
author: Brian B. Avants (<span style="color:blue;">PENN</span>) and <br />Nicholas J. Tustison (<span style="color:orange;">UVA</span>)
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

```{r global options, include=FALSE}
library(rmarkdown)
library(knitr)
figpath<-'figures/'
opts_chunk$set(fig.width=9, fig.height=6, fig.path=figpath,
               warning=FALSE, message=FALSE)
```

##

<div align="center"><img src="./logo.png" frameborder="0"></img></div>

This talk is online at [http://stnava.github.io/ANTs2015/](http://stnava.github.io/ANTs2015/) with colored [links](http://stnava.github.io/ANTs2015/) meant to be clicked for more information.

## ANTs & Neuroscience

We need statistical image analysis <br />at several scales in modern neuroscience

* Macro: *in vivo* structural and functional MRI

* Micro:  high-resolution post-mortem MRI links with in vivo MRI

* Nano: neuron reconstruction ...

* Solutions that are consistent across these scales have the potential
to build multi-scale feature sets or templates and provide new
insights into brain structure and function

* E.g. Parcellation constraints based on histology, tractography, function ...

* Statistical definitions of anatomy/pathology?

* Reinvention of these solutions within each lab ... can we mitigate this?

* Reduce, reuse, recycle ...

# *ANTs* optimizes mathematically <br /> well-defined  <br /><span style="color:red;">objective functions</span>  <br />guided by <br /> <span style="color:red;">prior knowledge</span>  ...

# plug *your ideas* <br />into our software: <br /> <span style="color:blue;"> <span style="color:red;">innovation, insight</span> <br />from biomedical data</span> ...

## Diffeomorphisms

<div align="center"><img width="1433" src="./figures/sillyputty.png" frameborder="0"></img></div>

plausible physical modeling of large, invertible deformations

"differentiable map with differentiable inverse"

## ANTs: Beyond Registration

![](./figures/antscapabilities.jpg)

[Atropos](http://www.ncbi.nlm.nih.gov/pubmed/?term=atropos+tustison) segmentation, [N4 inhomogeneity correction](http://www.ncbi.nlm.nih.gov/pubmed/?term=N4+tustison), [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/?term=eigenanatomy+avants), [SCCAN](http://www.ncbi.nlm.nih.gov/pubmed/?term=sparse+canonical+avants), [Prior-constrained PCA](http://www.ncbi.nlm.nih.gov/pubmed/24852460), and [atlas-based label fusion](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4009425/) and [MALF](http://www.ncbi.nlm.nih.gov/pubmed/21237273) (powerful expert systems for segmentation)

## Founding developers

<div align="center"><img width="1433" src="./figures/bant2.png" frameborder="0"></img></div>

## Long-term collaborators

![](./figures/antscollab.jpg)

$+$ [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

## General purpose library for multivariate image registration, segmentation & statistical analysis tools

* 170,000+ lines of C++, 6$+$ years of work, 15+ collaborators.

* Generic mathematical methods that are tunable for 
application specific domains:  no-free lunch

* Deep testing on multiple platforms ... osx, linux, windows.

* Several "wins" in public knock-abouts ( [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496), [Murphy 2011](http://www.ncbi.nlm.nih.gov/pubmed/21632295), [SATA 2012 and 2013](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837555/), [BRATS 2013](http://martinos.org/qtim/miccai2013/proc_brats_2013.pdf), others )

```
    An algorithm must use prior knowledge about a problem 
    to do well on that problem 
```

# <span style="color:red;">Medical Image Registration</span> <br /> Fundamental tool for<br /> morphometry, segmentation,<br /> motion estimation and <br /> <span style="color:blue;">data cleaning</span>

## Definitions

* Registration $=$ estimate an "optimal" geometric mapping between image pairs or image sets (e.g. Affine)

* Similarity $=$ a function relating one image to another, given a transformation (e.g. mutual information)

* <span style="color:grey;">
Diffeomorphisms $=$ differentiable map with differentiable inverse (e.g. "silly putty", viscous fluid)  </span>

* Segmentation $=$ labeling tissue or anatomy in images, usually automated (e.g. K-means)

* <span style="color:grey;"> Multivariate $=$ using many voxels or measurements at once (e.g. PCA, $p >> n$ ridge regression)</span>

* Multiple modality $=$ using many modalities at once (e.g. DTI and T1 and BOLD)

* MALF: multi-atlas label fusion - using anatomical dictionaries to label new data

* Solutions to challenging statistical image processing problems usually need elements from each of the above

## The Technical Framework

![](./figures/designprinciples.png)

... and most of it multivariate.

## [SyN](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2276735/) for optimization symmetry
![](./figures/fishSyN.png)
Images deform symmetrically along the shape manifold.  This eliminates bias in the
measurement of image differences.

##  [SyN (link)](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2276735/) Example

<!-- syn example -->

<div align="center"><iframe width="1220" height="730" src="http://www.youtube.com/embed/3I9RcRtpOvw" frameborder="0" allowfullscreen></iframe></div>

[SyN movie](http://www.youtube.com/embed/3I9RcRtpOvw)

## Concatenated transformation $+$ metric stages are necessary in real data

0.  Initialize the mapping ( more on this later )

1.  Start with a *rigid transformation*:  $I(x) \approx J(R(x))$ s.t. negative $MI$ is minimized

2.  Follow by an *affine transformation*:  $I(x) \approx J(R(A(x)))$ s.t. negative $MI$ is minimized with fixed $R$

3.  Finally, a *diffeomorphism*:  $I(x) \approx J(R(A(\phi(x))))$ s.t. $k$-neighborhood correlation $CC_k$ is minimized with fixed $R, A$

4.  Output the *composite transform* $A \circ R$ as a matrix transformation and $\phi$ and $\phi^{-1}$ as deformation fields.

standard in [recommended](http://stnava.github.io/ANTsDoc/) `antsRegistration` application scripts

## Registration benefits from<br /> *optimal sampling strategy*

* sampling for *both* the metric and the transformation

* impacts scalability, memory, optimization accuracy, speed, robustness ...

* could be done optimally *with massive improvements in performance*
    * but needs investment in order to achieve "dream" registration scenario

* important for new schemes that *elect* solutions from **anatomical or transformation** dictionaries

* overall, relatively little translational work on this important problem in biomedical imaging

## Sampling & feature selection: Multi-start

![](./figures/multistart.png)

Theoretical guarantee of global optimum: improves local optimizers.

Default in `antsCorticalThickness` pipeline and `FSL`.

## Sampling & feature selection: Biomedical imagery

![](./figures/slides1.png)

Initial configuration of data

## Sampling & feature selection: Biomedical imagery

![](./figures/slides2.png)

Automatic feature selection

## Sampling & feature selection: Biomedical imagery

![](./figures/slides3.png)

Resampling allows comparison & slide alignment and <br />
validates the feature selection

Dramatic reduction in computation time / memory requirements

## Sampling & feature selection: Lesioned brains

![](./figures/lesionedbrains.jpg)

## Sampling & feature selection: Summary

* we exploit these strategies to:
    - accelerate
    - focus 
    - validate

# Evaluation results

## 
![](./figures/evalhistory.png)

## Anatomical dictionaries
![](./figures/sata2013.png)

we provided the *standard* registration results for $>$ 20,000 image pairs at [SATA 2013](https://masi.vuse.vanderbilt.edu/workshop2013/index.php/MICCAI_2013_SATA_Challenge_and_Workshop:Current_events)

## [label fusion (link)](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3049832/)
![](./figures/sata2013.jpg)

## Multiple metrics improve performance
![](./figures/multivariateSATA.jpg)

to our knowledge, ANTs is the only freely available system that can solve this problem in a fully multivariate manner.

Hongzhi Wang won the "walk in the park" award for this work ...

# Differentiable maps with<br /> differentiable inverse <br />$+$ *statistics in these spaces*

## Brain templates as *high-dimensional averages*
![](./figures/speciestemplates.png)

## SyGN - templates and averages in deformation space

![](figures/normvsnorm.png)
from [miykael](http://miykael.github.io/nipype-beginner-s-guide/ANTS.html)

[geodesic image averages](http://www.ncbi.nlm.nih.gov/pubmed/15501083)

[optimal templates 2](http://www.ncbi.nlm.nih.gov/pubmed/19818860)

[canine template](http://www.ncbi.nlm.nih.gov/pubmed/23284904)

[chimpanzees](http://www.ncbi.nlm.nih.gov/pubmed/23516289)

## Statistics in deformation space

[AIBS movie](http://www.youtube.com/embed/8GgHG-rApiE)

## Average Republican and Democratic congressmen

<div align="center"><img src="figures/democrat_left_repub_right.png" frameborder="0"></img></div>

[congress](http://ntustison.github.io/CongressionalFaceTemplates/)

# ANTs versus Freesurfer:<br />  Quantifying *life span* brain health

## ANTs versus Freesurfer:<br />  Quantifying *life span* brain health

* Freesurfer is the historical standard for measuring cortical thickness

* instead of using surfaces to measure cortical thickness, we use the image space *DiReCTly*

* [see this section of a different talk](http://stnava.github.io/ANTsTalk/#/putting-it-all-together-can-we-quantify-life-span-brain-health-in-individuals-and-in-populations)

* and this "big data" paper: [Large-scale evaluation of ANTs and FreeSurfer cortical thickness measurements](http://www.ncbi.nlm.nih.gov/pubmed/24879923)

* comparison of prediction from automated cortical thickness measurement from 4 public datasets

* $>$ 1200 subjects, age 7 to over 90 years old

* *hint*: ANTs thickness measurements have higher prediction accuracy relative to Freesurfer ( implying we extract more information from the data )

* ANTs methods consistently improve statistical power [eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/24687814), [syn](http://www.ncbi.nlm.nih.gov/pubmed/?term=syn+epstein+avants), [itkv4](http://www.ncbi.nlm.nih.gov/pubmed/24817849) ... also, see [Schwarz CG, et al.](http://www.ncbi.nlm.nih.gov/pubmed/24650605) re: TBSS and related work in fMRI [Miller, PNAS](http://www.ncbi.nlm.nih.gov/pubmed/15980148), [Azab, et al in Hippocampus](http://www.ncbi.nlm.nih.gov/pubmed/24167043).

# Registration & statistics:<br /> Frontiers and innovation

## multivariate statistical fields arise from fused modalities

![images at measurement fields](figures/statisticalfields.png)

*Many opportunities for statistical advancements*

## Scientific Data 2014
![](./figures/ptbp2.png)

## "Network" of predictors for age

...

## ITK+ANTs+R = <span style="color:red;">*ANTsR*</span>

## Agnostic statistics
![](./figures/antsrex.jpg)

[McMillan et al.](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3961542/)

## A Quick <span style="color:grey;">*ANTsR*</span> example
```{r loadantsr,results='hide',echo=FALSE,message=FALSE}
library(ANTsR)
```

This is an executable *ANTsR* code block - *N*-dimensional statistics to go with 
our *N*-dimensional image processing software!
```{r loadimages,message=FALSE,warning=FALSE,results='hide'}
library(ANTsR)
dim<-2
filename<-getANTsRData('r16')
img<-antsImageRead( filename , dim )
filename<-getANTsRData('r64')
img2<-antsImageRead( filename , dim )
mask<-getMask(img,50,max(img),T)
mask2<-getMask(img,150,max(img),T)
nvox<-sum( mask == 1 )
nvox2<-sum( mask2 == 1 )
```
The brain has `r nvox` voxels ...

## A Quick <span style="color:grey;">*ANTsR*</span> example

Simulate a population morphometry study - a "VBM" ...
```{r morph,results='hide'}
simnum<-10
imglist<-list()
imglist2<-list()
for ( i in 1:simnum ) {
  img1sim<-antsImageClone(img)
  img1sim[ mask==1 ]<-rnorm(nvox,mean=0.5)
  img1sim[ mask2==1 ]<-rnorm(nvox2,mean=2.0)
  img2sim<-antsImageClone(img2)
  img2sim[ mask==1 ]<-rnorm(nvox,mean=0.20)
  imglist<-lappend(imglist,img1sim)
  imglist2<-lappend(imglist2,img2sim)
}
imglist<-lappend( imglist, imglist2 )
mat<-imageListToMatrix( imglist, mask )
DX<-factor( c( rep(0,simnum), rep(1,simnum) ) )
mylmresults<-bigLMStats( lm( mat ~ DX ) )
qvals<-p.adjust( mylmresults$pval.model ) 
```
The minimum q-value is `r min(qvals)` ...

## Visualize the histograms of effects
```{r vizmorph,results='hide'}
whichvox<-qvals < 1.e-2
voxdf<-data.frame( volume=c( as.numeric( mat[,whichvox] ) ), DX=DX )
ggplot(voxdf, aes(volume, fill = DX)) + geom_density(alpha = 0.2)
```

## Visualize the anatomical distribution
```{r vizmorph2hide,results='hide',echo=FALSE}
betas<-antsImageClone(mask)
betavec<-abs(mylmresults$beta.t[1,])
betas[mask==1]<-betavec
SmoothImage(dim,betas,1.5,betas)
thresh<-paste(3.0,'x',max(betavec),sep='')
ofn<-paste(figpath,'vizmorph2.jpg',sep='')
```

```{r vizmorph2,results='hide'}
plotANTsImage(img,functional=list(betas),threshold=thresh,
  outname=ofn)
```

![](`r ofn`)

## Network visualization
see `?plotBasicNetwork`
```{r networkfn,results='hide',warning=FALSE,echo=FALSE,message=FALSE}
ofn<-paste(figpath,'network',sep='')
ofnpng<-paste(ofn,'.png',sep='')
```
```{r network,results='hide',warning=FALSE,echo=FALSE,message=FALSE,eval=FALSE}
       mnit<-getANTsRData("mni")
       mnit<-antsImageRead(mnit,3)
       mnia<-getANTsRData("mnia")
       mnia<-antsImageRead(mnia,3)
       ThresholdImage(3,mnit,mnit,1,max(mnit))
       ImageMath(3,mnit,"FillHoles",mnit)
       cnt<-getCentroids( mnia, clustparam = 50 )
       aalcnt<-cnt$centroids[1:90,]
       brain<-renderSurfaceFunction( surfimg =list( mnit ) , alphasurf=0.1 ,smoothsval = 1.5 )
       testweights<-matrix( rep( 0, 90*90 ) ,nrow=90)
       testweights[31,37]<-1  # ant cingulate to hipp
       testweights[31,36]<-2  # ant cingulate to post cingulate
       testweights[11,65]<-3  # broca to angular 
       plotBasicNetwork( centroids = aalcnt , brain , weights=testweights )
       id<-par3d("userMatrix")
       rid<-rotate3d( id , -pi/2, 1, 0, 0 )
       rid2<-rotate3d( id , pi/2, 0, 0, 1 )
       rid3<-rotate3d( id , -pi/2, 0, 0, 1 )
       par3d(userMatrix = id ) 
       dd<-make3ViewPNG(  rid, id, rid2,  ofn )
       par3d(userMatrix = id ) 
```
![](`r ofnpng`)

## The power of *ANTs* $+$ *R* $\rightarrow$<br /> <span style="color:red;">**Reproducible imaging science**</span>

![](./figures/antsgoals.png)

... used in ["Sparse canonical correlation analysis relates network-level atrophy to multivariate cognitive measures in a neurodegenerative population"](http://www.ncbi.nlm.nih.gov/pubmed/24096125) and several upcoming ...

## Can we customize these methods for a challenging multivariate segmentation problem with clinical relevance?

## Automated tumor segmentation from multiple MRI
![](./figures/tumorsegmentation.png)

## General theory tunable to specific domains:  *no-free lunch*
![](./figures/brats.png)

[BRATS 2013 Challenge](http://hal.inria.fr/docs/00/93/56/40/PDF/brats_preprint.pdf) **won** with [*ANTsR*](https://github.com/stnava/ANTsR)

## Can we address subtle questions in brain and cognition via imaging-specific dimensionality reduction?

##  [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/24852460) & [SCCAN](http://www.ncbi.nlm.nih.gov/pubmed/?term=avants+SCCAN)

## ANTs*R* $\rightarrow$ new insight <br />via better analytics

Good software should fade into the background ... however ...

<span style="color:red;">
 As is common in science, the first big breakthrough in our understanding ...
 [came from] an improvement in measurement.
</span>

<span style="color:blue;">
> Daniel Kahnemann, *Thinking, Fast and Slow* (2011)
</span>

# Conclusion

## Recap

- Powerful, general-purpose, <span style="color:red;">well-evaluated</span> registration and segmentation.

- Differentiable maps with differentiable inverse <span style="color:red;">$+$ statistics in these spaces</span>

- Evaluated in multiple problem domains</span> via internal studies & open competition

- Borg philosophy: <span style="color:red;">"best of"</span> from I/O, to processing to statistical methods

- Open source, testing, many examples, consistent style, multiple platforms,  active community support ...

- Integration with *R* $+$ novel tools for prediction, decoding, high-to-low dimensional statistics.

- Collaborations with [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

## Challenges: Computational and Scientific

- Scalability
    - **need to fuse feature selection methods with transformation optimization**
    - **need to leverage existing ITK streaming infrastructure in application level tool**
- Domain expertise: Customizable for specific problems but sometimes not specific enough
- "Plausible physical modeling ..." - this should vary per problem ... but doesn't.
    - a fabulous project would be to resolve this issue at a large-scale e.g. for reconstructing neurons, measuring white matter elaboration ...
    - our prior FEM work is one potential solution
- Rapid development: colleagues still need familiarity with compilation for latest ANTs features
- Latest theoretical advances in registration not yet wrapped for users
- Need more [Documentation](http://stnava.github.io/ANTs/)  & [testing](http://testing.psychiatry.uiowa.edu/CDash/index.php?project=ANTS) ...

## Tools you can use for imaging science

- Core developers:  *B. Avants, N. Tustison, H. J. Johnson, J. T. Duda*

- Many contributors, including users ...

- Multi-platform, multi-threaded C++ [stnava.github.io/ANTs](stnava.github.io/ANTs)

- Developed in conjunction with [http://www.itk.org/](http://www.itk.org/)

- R wrapping and extension [stnava.github.io/ANTsR](stnava.github.io/ANTsR)

- rapid development, regular testing $+$ many eyes $\rightarrow$ bugs are shallow

<img style="float: right" src="figures/penn.png" />
<img style="float: left" src="figures/picsl.jpg" />

# Analysis philosophy and <br /> published opinions

## What is and *is not*<br /> image registration

[*not* registration](http://www.ncbi.nlm.nih.gov/pubmed/23116330)

## Voodoo in voxel-based analysis

[logical circularity VBA](http://www.ncbi.nlm.nih.gov/pubmed/?term=logical+circularity+tustison)

## Instrumentation bias in the use and evaluation of software

[Instrumentation bias in the use and evaluation of software](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3766821/?report=reader)

# References {#reffont}