### 2014-09-02 lesions/lesions.Rmd
ENH:  Adding lesion stuff.

# Cortical thickness with lesions

## antsCorticalThickness.sh is adaptable!

1. Register subject (or single subject template) to normal template.

2. Transform lesion mask to normal template.

3. Create additional "lesion" prior, i.e. ```SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1```.

4.  Subtract out lesion prior from all other priors and keep values $\in [0,1]$.

## Modified template spatial priors

![](lesions/figures/lesionPrior.png)

## antsCorticalThickness.sh using lesion prior

![](lesions/figures/lesionSegmentation.png)

Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."