### 2014-09-03 background/background1.Rmd
WIP: regular update

# Background

## Founding developers

<div align="center"><img width="1800" src="./figures/bant2.png" frameborder="0"></img></div>

BBA & NT

## Long-term collaborators

![](./figures/antscollab.jpg)

$+$ [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

##

![](./figures/lion.png)

## Definitions

* Registration $=$ estimate an "optimal" geometric mapping between image pairs or image sets (e.g. Affine)

* Similarity $=$ a function relating one image to another, given a transformation (e.g. mutual information)

* <span style="color:grey;">
Diffeomorphisms $=$ differentiable map with differentiable inverse (e.g. "silly putty", viscous fluid)  </span>

* Segmentation $=$ labeling tissue or anatomy in images, usually automated (e.g. K-means)

* <span style="color:grey;"> Multivariate $=$ using many voxels or measurements at once (e.g. PCA, $p >> n$ ridge regression)</span>

* Multiple modality $=$ using many modalities at once (e.g. DTI and T1 and BOLD)

* MALF: multi-atlas label fusion - using anatomical dictionaries to label new data

* Solutions to challenging statistical image processing problems usually need elements from each of the above

## Image mapping & perception: 1878

![](./figures/galton.png)

* Francis Galton: Can we see criminality in the face?

* (maybe he should have used ANTs?)

## Initial scope

![](./figures/ants_initial_scope.png)

## ANTs Lineage

![](./figures/lineage.jpg)

References: @Horn1981, @Gee1993, @Grenander1993, @Thompson2001, @Miller2002, @Shen2002, @Arnold2014, @Thirion1998, @Rueckert1999, @Fischl2012, @Ashburner2012

## Diffeomorphisms

<div align="center"><img width="1433" src="./figures/sillyputty.png" frameborder="0"></img></div>

plausible physical modeling of large, invertible deformations

"differentiable map with differentiable inverse"

## Fine-grained and flexible maps
![](./figures/highresdiffeos.jpg)

... to correct a misconception about diffeomorphisms ...

## General purpose library for multivariate image registration, segmentation & statistical analysis tools

* 170,000+ lines of C++, 6$+$ years of work, 15+ collaborators.

* Generic mathematical methods that are tunable for 
application specific domains:  no-free lunch

* Deep testing on multiple platforms ... osx, linux, windows.

* Several "wins" in public knock-abouts ( [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496), [Murphy 2011](http://www.ncbi.nlm.nih.gov/pubmed/21632295), [SATA 2012 and 2013](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837555/), [BRATS 2013](http://martinos.org/qtim/miccai2013/proc_brats_2013.pdf), others )

```
    An algorithm must use prior knowledge about a problem 
    to do well on that problem

## ANTs: Beyond Registration

![](./figures/antscapabilities.jpg)

[Atropos](http://www.ncbi.nlm.nih.gov/pubmed/?term=atropos+tustison) segmentation, [N4 inhomogeneity correction](http://www.ncbi.nlm.nih.gov/pubmed/?term=N4+tustison), [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/?term=eigenanatomy+avants), [SCCAN](http://www.ncbi.nlm.nih.gov/pubmed/?term=sparse+canonical+avants), [Prior-constrained PCA](http://www.ncbi.nlm.nih.gov/pubmed/24852460), and [atlas-based label fusion](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4009425/) and [MALF](http://www.ncbi.nlm.nih.gov/pubmed/21237273) (powerful expert systems for segmentation)

```

## On documentation

![](figures/handsaw.jpg)

simple and nearly self-explanatory

## On documentation

![](figures/Chainsaw.jpg)

documentation is important

## On documentation

![](figures/blindmen.jpg)