### 2014-09-03 conclusion/conclusion.Rmd
WIP: regular update

## Questions driving *ANTs* ($+$ Refs.1)

* how should we geometrically transform anatomical coordinates?

    - [syn paper](http://www.ncbi.nlm.nih.gov/pubmed/17659998) - our geometric transformation model of choice
    - [recent B-spline alternative/improvement](http://www.ncbi.nlm.nih.gov/pubmed/24409140)

* how should we measure pairwise image similarity?

    - [similarity metric evaluation](http://www.ncbi.nlm.nih.gov/pubmed/20851191) compares functions for computing rigid or affine transformations between images

* what if this pair has rgb/vector/tensor voxels?

    - [dti similarity](http://www.ncbi.nlm.nih.gov/pubmed/18041273)

* how do we extend from pairs to hundreds or thousands of pairs of images?

    - [optimal templates](http://www.ncbi.nlm.nih.gov/pubmed/?term=avants+optimal+template) and [spatial priors](http://www.ncbi.nlm.nih.gov/pubmed/24879923)
    - species specific templates/priors in [chimps](http://www.ncbi.nlm.nih.gov/pubmed/23516289) and [canines](http://www.ncbi.nlm.nih.gov/pubmed/23284904)
    - [ants "big data"](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+freesurfer)

* how do we fuse multiple modality images at the subject and population levels?

    - [ants auxiliary modality study](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+logical+circularity)

## Questions driving *ANTs* ($+$ Refs.2)

* can diffeomorphisms [improve cortical thickness measurement](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+freesurfer)?

* how might we efficiently cluster the statistical fields that arise in image analysis?

    - [Atropos](http://www.ncbi.nlm.nih.gov/pubmed/?term=atropos+tustison) segmentation and [N4 inhomogeneity correction](http://www.ncbi.nlm.nih.gov/pubmed/?term=N4+tustison)
    - [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/?term=eigenanatomy+avants) for sparse imaging-specific PCA

* how to cluster such fields when we have supervision?

    - [sparse canonical correlation analysis for neuroimaging](http://www.ncbi.nlm.nih.gov/pubmed/?term=sparse+canonical+avants)
    - [Prior-constrained PCA](http://www.ncbi.nlm.nih.gov/pubmed/24852460)
    - [atlas-based label fusion](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4009425/) and [MALF](http://www.ncbi.nlm.nih.gov/pubmed/21237273) - powerful expert systems for segmentation

* how do we implement a fully multivariate *interpretable* brain and behavior study?

    - [SCCAN for imaging & cognition](http://www.ncbi.nlm.nih.gov/pubmed/?term=sccan+avants)

* how do we extend these ideas to functional MRI & decoding?

    - [recent unpublished software](http://stnava.github.io/RKRNS/) 
    - [recent work with Ben Kandel](..not yet...)

## *ANTs* longitudinal analysis

* Longitudinal image processing issues
    - [registration induced bias](http://www.ncbi.nlm.nih.gov/pubmed/20005963)
    - [general & TBI-specific issues in longitudinal analysis](http://www.ncbi.nlm.nih.gov/pubmed/23549059)
    - [reproducibility of CBF](http://www.ncbi.nlm.nih.gov/pubmed/22517961)
    - [structure-specific analysis](http://www.ncbi.nlm.nih.gov/pubmed/22306801)

* An early study of longitudinal cortical change in ALS
    - [ALS atrophy rates](http://www.ncbi.nlm.nih.gov/pubmed/16317254)

* Extension of standard-setting ants cortical thickness pipeline to longitudinal data
    - [DynANTs (unpublished)](https://github.com/stnava/DynANTs)

## Challenges: Computational and Scientific

- Scalability: **need to fuse feature selection methods with transformation optimization**
- Scalability: **need to leverage existing ITK streaming infrastructure in application level tool**
- Domain expertise: Customizable for specific problems but sometimes not specific enough
- Rapid development: colleagues still need familiarity with compilation for latest ANTs features
- Latest theoretical advances in registration not yet wrapped for users
- Need more [Documentation](http://stnava.github.io/ANTs/)  & [testing](http://testing.psychiatry.uiowa.edu/CDash/index.php?project=ANTS) ...

## Recap

- Powerful, general-purpose, <span style="color:red;">well-evaluated</span> registration and segmentation.

- Differentiable maps with differentiable inverse <span style="color:red;">$+$ statistics in these spaces</span>

- Evaluated in multiple problem domains</span> via internal studies & open competition

- Borg philosophy: <span style="color:red;">"best of"</span> from I/O, to processing to statistical methods

- Open source, testing, many examples, consistent style, multiple platforms,  active community support ...

- Integration with *R* $+$ novel tools for prediction, decoding, high-to-low dimensional statistics.

- Collaborations with [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

## Tools you can use for imaging science

- Core developers:  *B. Avants, N. Tustison, H. J. Johnson, J. T. Duda*

- Many contributors, including users ...

- Multi-platform, multi-threaded C++ [stnava.github.io/ANTs](stnava.github.io/ANTs)

- Developed in conjunction with [http://www.itk.org/](http://www.itk.org/)

- R wrapping and extension [stnava.github.io/ANTsR](stnava.github.io/ANTsR)

- rapid development, regular testing $+$ many eyes $\rightarrow$ bugs are shallow

<img style="float: right" src="figures/penn.png" />
<img style="float: left" src="figures/picsl.jpg" />