### 2014-09-03 impact/impact1.Rmd
WIP: regular update

# ANTs Impact

## ANTs Statistics

![](impact/figures/antsvisitors.png)

## ANTs & Neuroscience

We need statistical image analysis <br />at several scales in modern neuroscience

* Macro: *in vivo* structural and functional MRI

* Micro:  high-resolution post-mortem MRI links with in vivo MRI

* Nano: neuron reconstruction ...

* Solutions that are consistent across these scales have the potential
to build multi-scale feature sets or templates and provide new
insights into brain structure and function

* E.g. Parcellation constraints based on histology, tractography, function ...

* Statistical definitions of anatomy/pathology?

* Reinvention of these solutions within each lab ... can we mitigate this?

* Reduce, reuse, recycle ...

## Financial impact

* The **high error rate** in software makes testing for quality assurance critically important.

* Bugs exist **Everywhere**: so don't freak out if we admit we have bugs:  we're just being honest ( and not all devs are )

* Virtually every business in the United States now depends on software for development, production, distribution, and after-sales support of products and services.

* A 2002 NIST study estimated the direct costs to the software supply chain due to failure to identify (**successfully test for**) "bugs"

* We are better at this than most ... **not perfect**

* The estimate of direct costs compiled from industry survey data for the U.S. economy was **\$60 billion per year**

* ... this estimate did not include costs to end users such as lost business (for example, the cost of shutting down the New York Mercantile Exchange in 1998 due to a software failure).

## Reproducibility

* Consider the recent paper: [Freesurfer, laplacian, DiReCT thickness in longitudinal stroke](http://ac.els-cdn.com/S2213158214001260/1-s2.0-S2213158214001260-main.pdf?_tid=8e6f5248-335d-11e4-ba4f-00000aab0f02&acdnat=1409743944_6615d59eb5709e6bcb1aca76450f7d93)

* We "pre-wrote" a comment: [Instrumentation bias in the use and evaluation of scientific software](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3766821/?report=reader)

* With these guys:

* <img src="figures/luis.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right: 3px;"><img src="figures/torsten.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;">

## Hired by Google

<img src="figures/luis.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right: 3px;"><img src="figures/torsten.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;"><img src="figures/gang.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;">