### 2014-09-03 longitudinal/longitudinal.Rmd
WIP: regular update

# New work

## RKRNS

[link](stnava.github.io/RKRNS)

What mathematical encoding of language reflects both brain activity and 
"our" understanding of semantics?

## RIPMARC: Imputation, strutural networks, etc.

![](figures/ripmmarc_inter.png)

MICCAI 2014, Neuroimage 2015

## Big image registration

# Longitudinal processing with ANTs

## ANTs longitudinal pipeline

![](longitudinal/figures/longpipe.png)

see [link to paper](http://link.springer.com/chapter/10.1007/978-3-642-15705-9_40)

and [unbiased analysis paper](http://www.ncbi.nlm.nih.gov/pubmed/20005963)

plus [longitudinal recommendation paper](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3581852/)

## ANTs longitudinal pipeline

![](longitudinal/figures/longpipeex.png)

## ANTs longitudinal pipeline: TBI

![](longitudinal/figures/long_with_lesions_seg.png)

## ANTs longitudinal pipeline: TBI

![](longitudinal/figures/long_with_lesions_thick.png)

## ANTs longitudinal pipeline

![](longitudinal/figures/babystudy.png)