### 2014-10-15 tools/n4/n4.Rmd
ENH:  Adding tools subfolder.

# N4

## Nonparametric nonuniform intensity normalization (N3)

* Developed at the Montreal Neurological Institute (John Sled, 1998)

* Part of the standard preprocessing protocol in large scale projects such as ADNI

* The traditional de facto standard in MRI bias correction
    + good performance
    + *public availability*

* Public availability --- set of perl scripts coordinating various C++ programs

* "*Let's incorporate N3 into ANTs!*"

## N3 adoption issues

![](./tools/n4/figures/whyN4.png)

<!-- regular html comment

<div align="center"><img src="./tools/n4/figures/whyN4.png" frameborder="0"></img></div>

-->

"*So let's re-implement N3 in ITK for ANTs users.*"

## N3 -> N4<sup>*</sup>

<p align = "middle"><small><sup>*</sup><u>N</u>ick's N3</small></p>

* comparative [evaluation](http://www.ncbi.nlm.nih.gov/pubmed/20378467)

* smaller spline distances (useful for higher magnet strengths)

* multiresolution

* weighted regional mask (used in [`antsAtroposN4.sh`](https://github.com/stnava/ANTs/blob/master/Scripts/antsAtroposN4.sh))

* faster execution times

* less susceptible to noise

* _publicly available_

## N4 example

![](./tools/n4/figures/KKI2009_MPRAGE_slice150.png)

## Command line help

<font size="5", color="lightgrey">
```
$ N4BiasFieldCorrection -h

COMMAND:
     N4BiasFieldCorrection

OPTIONS:
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -r, --rescale-intensities 0/(1)
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50x50x50x50>,<convergenceThreshold=0.0>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>]
                           [initialMeshResolution,<splineOrder=3>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output correctedImage
                  [correctedImage,<biasField>]
     -h
     --help
```
</font>