### 2014-10-16 competitions/competitions.Rmd

# Competitions

##

![](./evaluation/figures/evalhistory.png)

## <span style="color:green;">BRATS 2013</span> <small>brain tumor segmentation</small>

![](./competitions/figures/brats2013results.png)

_"Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with ANTsR"_

## <span style="color:green;">STACOM 2014</span> <small>cardiac motion estimation</small>

![](./competitions/figures/cardiacBSplineSyN.png)

A great opportunity to showcase a [new SyN variant](http://www.ncbi.nlm.nih.gov/pubmed/24409140)