### 2014-10-17 ants2015_GE.Rmd
ENH:  Refinements.

---
title: <span style="color:red;">ANTs</span><br> <small>_power tools for neuroimaging_</small>
author:
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

```{r global options, include=FALSE}
library(rmarkdown)
library(knitr)
figpath<-'figures/'
opts_chunk$set(fig.width=9, fig.height=6, fig.path=figpath,
               warning=FALSE, message=FALSE)
```

# Background

##

![](./figures/brian_and_nick.png)

## Built on ITK

* image registration framework

* many other classes such as:
    + `itkScalarImageToRGBImageFilter`
    + `itkLabelOverlapMeauresImageFilter`
    + `itkBSplineScatteredDataPointSetToImageFilter`
    + `itkStochasticFractalDimensionImageFilter`

## ANTs development philosophy:<br><small>Assimilate the optimal components of other approaches</small>

![](./figures/BorgFirstContact.jpg)

_"We are the Borg. Your biological and technological distinctiveness will be added to our own. Resistance is futile."_

## *The Big 5*

* <span style="color:red">image registration</span> (`antsRegistration.cxx`)

* <span style="color:red">bias correction</span> (`N4BiasFieldCorrection.cxx`)

* multivariate template construction (`antsMultivariateTemplateConstruction.sh`)

* multivariate segmentation (`Atropos.cxx`)

* cortical thickness (`KellyKaposki.cxx` and `antsCorticalThickness.sh`)

## *ANTsR*<br>for statistics and visualization

# *ANTs* optimizes mathematically well-defined <span style="color:red;">objective functions</span> guided by <span style="color:red;">prior knowledge</span>  ...

# ... including that of developers, domain experts and other colleagues ...

# plug *your ideas* into our software to gain insight into biomedical data ...

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains