### 2014-10-23 ants2015build.Rmd
ENH:  Adding more circularity material.

---
title: <span style="color:red;">*ANTs*</span><br> <small>_power tools for neuroimaging_</small>
author:
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

```{r global options, include=FALSE}
library(rmarkdown)
library(knitr)
figpath<-'figures/'
opts_chunk$set(fig.width=9, fig.height=6, fig.path=figpath,
               warning=FALSE, message=FALSE)
```

# Background

##

![](./figures/brian_and_nick.png)

## Built on ITK

* image registration framework

* many other classes such as:
    + `itkScalarImageToRGBImageFilter`
    + `itkLabelOverlapMeauresImageFilter`
    + `itkBSplineScatteredDataPointSetToImageFilter`
    + `itkStochasticFractalDimensionImageFilter`

## ANTs development philosophy:<br><small>Assimilate the optimal components of other approaches</small>

![](./figures/BorgFirstContact.jpg)

_"We are the Borg. Your biological and technological distinctiveness will be added to our own. Resistance is futile."_

## *The Big 5*

* <span style="color:red">image registration</span> (`antsRegistration.cxx`)

* <span style="color:red">bias correction</span> (`N4BiasFieldCorrection.cxx`)

* multivariate template construction (`antsMultivariateTemplateConstruction.sh`)

* multivariate segmentation (`Atropos.cxx`)

* cortical thickness (`KellyKaposki.cxx` and `antsCorticalThickness.sh`)

## *ANTsR*<br>for statistics and visualization

# *ANTs* optimizes mathematically well-defined <span style="color:red;">objective functions</span> guided by <span style="color:red;">prior knowledge</span>  ...

# ... including that of developers, domain experts and other colleagues ...

# plug *your ideas* into our software to gain insight into biomedical data ...

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains

# *ANTs* and the perils of circularity

## *ANTs* and the perils of circularity

![](./philosophy/circularity/figures/escherANTs.png)

##

![](./philosophy/circularity/figures/voodoo.png)

##

![](./philosophy/circularity/figures/fastForward.png)

##

![](./philosophy/circularity/figures/tbss.png)

##

![](./philosophy/circularity/figures/antsFlavoredTbss1.png)

##

![](./philosophy/circularity/figures/antsFlavoredTbss2.png)

##

![](./philosophy/circularity/figures/antsFlavoredTbss3.png)

##

![](./philosophy/circularity/figures/antsFlavoredTbss4.png)

##

![](./philosophy/circularity/figures/antsFlavoredTbss5.png)

##

![](./philosophy/circularity/figures/brianEmail.png)

##

![](./philosophy/circularity/figures/sleeping1.png)

##

![](./philosophy/circularity/figures/sleeping2.png)

##

![](./philosophy/circularity/figures/sleeping0.png)

##

![](./philosophy/circularity/figures/sleeping4.png)

##

![](./philosophy/circularity/figures/sleeping5.png)

##

![](./philosophy/circularity/figures/sleeping6.png)

##

![](./philosophy/circularity/figures/sleeping7.png)

##

![](./philosophy/circularity/figures/benefitsOfOpenScience.png)

## Command line help

<font size="5", color="lightgrey">
```
$ CreateDTICohort -h

COMMAND:
     CreateDTICohort

OPTIONS:
     -d, --image-dimensionality 2/3
     -a, --dti-atlas inputDTIAtlasFileName
     -x, --label-mask-image maskImageFileName
                            lowerThresholdFunction
     -n, --noise-sigma <noiseSigma=18>
     -p, --pathology label[<percentageChangeEig1=-0.05>,<percentageChangeAvgEig2andEig3=0.05>,<numberOfVoxels=all or percentageOfvoxels>]
     -w, --dwi-parameters [B0Image,directionFile,bvalue]
                          [B0Image,schemeFile]
     -r, --registered-population textFileWithFileNames.txt
     -o, --output [outputDirectory,fileNameSeriesRootName,<numberOfControls=10>,<numberOfExperimentals=10>]
     -h
     --help
```
</font>

## Which one is not warped?

![](./philosophy/circularity/figures/whichOneIsRegistered.png)

## _Answer_

![](./philosophy/circularity/figures/whichOneIsRegistered2.png)

##

![](./philosophy/circularity/figures/blobs1.png)

##

![](./philosophy/circularity/figures/blobs2.png)

##

![](./philosophy/circularity/figures/BooksteinReview1.png)
![](./philosophy/circularity/figures/BooksteinReview2.png)
![](./philosophy/circularity/figures/BooksteinReview3.png)