### 2015-04-28 competitions/competitions.Rmd
ENH:  More slides for compeittion.

# Competitions

## <span style="color:green;">Competitions</span>

* [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496)
* [EMPIRE 2010](http://empire10.isi.uu.nl)
* [Multi-Atlas Label Challenge 2012](https://masi.vuse.vanderbilt.edu/workshop2012/index.php/Main_Page)
* [SATA Challenge 2013](https://masi.vuse.vanderbilt.edu/workshop2013/index.php/MICCAI_2013_SATA_Challenge_and_Workshop:Current_events)
* [BRATS 2013](http://martinos.org/qtim/miccai2013/)
* [STACOM 2014 MoCo Challenge](http://www.cardiacatlas.org/web/stacom2014/moco-introduction)

<!--- ![](./evaluation/figures/evalhistory.png) -->

## <span style="color:green;">Klein 2009</span> <small>brain image registration</small>

![](./competitions/figures/nihms-130679-f0001.jpg)

* 14 algorithms
* publicly available data
* unbiased evaluation

## <span style="color:green;">Klein 2009</span> <small>results</small>

![](./competitions/figures/KleinResults2009.png)

_"One of the most significant findings of this study is that the relative performances of
the registration methods under comparison appear to be little affected by the choice of
subject population, labeling protocol, and type of overlap measure....  ART, SyN, IRTK, and
SPM's DARTEL Toolbox gave the best results according to overlap and distance measures,
with ART and SyN delivering the most consistently high accuracy across subjects and label sets."_

## <span style="color:green;">EMPIRE 2010</span> <small>lung image registration</small>

## <span style="color:green;">Multi-Atlas Label Challenge 2012</span> <small>atlas label fusion</small>

## <span style="color:green;">SATA Challenge 2012</span> <small>multi-modal/multi application label fusion</small>

## <span style="color:green;">BRATS 2013</span> <small>brain tumor segmentation</small>

![](./competitions/figures/brats2013results.png)

_Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with ANTsR_

## <span style="color:green;">STACOM 2014</span> <small>cardiac motion estimation</small>

![](./competitions/figures/cardiacBSplineSyN.png)

A great opportunity to showcase a [new SyN variant](http://www.ncbi.nlm.nih.gov/pubmed/24409140)