### 2015-04-29 competitions/competitions.Rmd
ENH:  More comp. stuff.

# Competitions

## <span style="color:green;">Competitions</span>

* [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496)
* [EMPIRE 2010](http://empire10.isi.uu.nl)
* [Multi-Atlas Label Challenge 2012](https://masi.vuse.vanderbilt.edu/workshop2012/index.php/Main_Page)
* [SATA Challenge 2013](https://masi.vuse.vanderbilt.edu/workshop2013/index.php/MICCAI_2013_SATA_Challenge_and_Workshop:Current_events)
* [BRATS 2013](http://martinos.org/qtim/miccai2013/)
* [STACOM 2014 MoCo Challenge](http://www.cardiacatlas.org/web/stacom2014/moco-introduction)

<!--- ![](./evaluation/figures/evalhistory.png) -->

## <span style="color:green;">Klein 2009</span>

![](./competitions/figures/nihms-130679-f0001.jpg)

* 14 algorithms
* publicly available data
* different labeling protocols
* developers tune their own algorithms

## <span style="color:green;">Results</span>

![](./competitions/figures/KleinResults2009.png)

_"One of the most significant findings of this study is that the relative performances of
the registration methods under comparison appear to be little affected by the choice of
subject population, labeling protocol, and type of overlap measure....  ART, SyN, IRTK, and
SPM's DARTEL Toolbox gave the best results according to overlap and distance measures,
with ART and SyN delivering the most consistently high accuracy across subjects and label sets."_

## <span style="color:green;">EMPIRE 2010</span>

![](./competitions/figures/EMPIRE10criteria.png)

* Alignment of lung boundaries,
* major fissures,
* annotated landmark pairs, and
* topology of displacement field.

## <span style="color:green;"> Results</span>

![](./competitions/figures/EMPIRE10Results.png)

## <span style="color:green;">Multi-Atlas 2012</span>

![](./competitions/figures/NKI30_MALF.png)

## <span style="color:green;">Results</span>

![](./competitions/figures/MALF2012.png)

## <span style="color:green;">SATA 2013</span>

## <span style="color:green;">BRATS 2013</span>

![](./competitions/figures/brats2013results.png)

_Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with ANTsR_

## <span style="color:green;">STACOM 2014</span>

![](./competitions/figures/cardiacBSplineSyN.png)

A great opportunity to showcase a [new SyN variant](http://www.ncbi.nlm.nih.gov/pubmed/24409140)