### 2015-05-08 installation/ants_installation.Rmd
ENH:  Installation and initial testing.

# ANTs installation

##

ANTs is a <span style="color:red;">_dynamic_</span> toolkit so,

![](./installation/figures/coder2.png)

for the willing, we recommend compiling from source.

## Requirements for source installation

* [CMake](http://www.cmake.org) -- build system
* [git](http://git-scm.com/downloads) -- version control system
* C++ compiler
* ITK/VTK dependencies are handled by CMake magic

##

https://github.com/stnava/ANTs

* README.md

    * "Is the current code building without compilation errors?"
    * More information about ANTs.

* Where to report [issues](https://github.com/stnava/ANTs/issues)

* or [contribute](https://github.com/stnava/ANTs/pulls).

## From the terminal...

~~~~~~~~~~~~~~~~~~
$ cd ${WHICH_PATH}
$ git clone https://github.com/stnava/ANTs.git
$ cd ANTs
$ mkdir bin
$ cd bin
$ ccmake ..
$ make
$ echo "export ANTSPATH=/${WHICH_PATH}/ANTs/bin/bin/" >> ~/.profile
$ echo "export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=4" >> ~/.profile
$ source ~/.profile
~~~~~~~~~~~~~~~~~~

##

<span style="color:blue;">__Some cmake considerations:__</span>

* Installation location?  (default = /usr/local/bin)
* Enable external applications (only ```jointFusion```)
* Type of build (a debug build produces __slow__ programs)

## Now test installation

##

Try calling the <span style="color:blue;">```KellyKapowski```</span> short help menu.

~~~~~~~~~~~~~~~~~~
$ KellyKapowski -h

COMMAND:
     KellyKapowski

OPTIONS:
     -d, --image-dimensionality 2/3
     -s, --segmentation-image imageFilename
                              [imageFilename,<grayMatterLabel=2>,<whiteMatterLabel=3>]
     -g, --gray-matter-probability-image imageFilename
     -w, --white-matter-probability-image imageFilename
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>,<convergenceWindowSize=10>]
     -t, --thickness-prior-estimate thicknessPriorEstimate
     -a, --thickness-prior-image thicknessPriorFileName
     -r, --gradient-step stepSize
     -l, --smoothing-variance variance
     -m, --smoothing-velocity-field-parameter variance
                                              isotropicMeshSpacing
     -b, --use-bspline-smoothing 1/(0)
     -n, --number-of-integration-points numberOfPoints
     -p, --maximum-number-of-invert-displacement-field-iterations numberOfIterations
     -o, --output imageFileName
                  [imageFileName,warpedWhiteMatterImageFileName]
     -v, --verbose (0)/1
     -h
     --help
~~~~~~~~~~~~~~~~~~

##

or the <span style="color:blue;">```antsRegistration```</span> long help menu.

~~~~~~~~~~~~~~~~~~
$ antsRegistration --help

COMMAND:
     antsRegistration
          This program is a user-level registration application meant to utilize
          ITKv4-only classes. The user can specify any number of "stages" where a stage
          consists of a transform; an image metric; and iterations, shrink factors, and
          smoothing sigmas for each level.Note that dimensionality, metric, transform,
          output, convergence, shrink-factors and smoothing-sigmas parameters are
          mandatory.

OPTIONS:
     --version
          Get Version Information.

     -d, --dimensionality 2/3
          This option forces the image to be treated as a specified-dimensional image. If
          not specified, we try to infer the dimensionality from the input image.

     -o, --output outputTransformPrefix
                  [outputTransformPrefix,<outputWarpedImage>,<outputInverseWarpedImage>]
          Specify the output transform prefix (output format is .nii.gz ). Optionally, one
          can choose to warp the moving image to the fixed space and, if the inverse
          transform exists, one can also output the warped fixed image. Note that only the
          images specified in the first metric call are warped. Use antsApplyTransforms to
          warp other images using the resultant transform(s).

     -j, --save-state saveSateAsTransform
          Specify the output file for the current state of the registration. The state
          file is written to an hdf5 composite file. It is specially usefull if we want to
          save the current state of a SyN registration to the disk, so we can load and
          restore that later to continue the next registration process directly started
          from the last saved state. The output file of this flag is the same as the
          write-composite-transform, unless the last transform is a SyN transform. In that
          case, the inverse displacement field of the SyN transform is also added to the
          output composite transform. Again notice that this file cannot be treated as a
          transform, and restore-state option must be used to load the written file by
          this flag.

     -k, --restore-state restoreStateAsATransform
          Specify the initial state of the registration which get immediately used to
          directly initialize the registration process. The flag is mutually exclusive
          with other intialization flags.If this flag is used, none of the
          initial-moving-transform and initial-fixed-transform cannot be used.

     -a, --write-composite-transform 1/(0)
          Boolean specifying whether or not the composite transform (and its inverse, if
          it exists) should be written to an hdf5 composite file. This is false by default
          so that only the transform for each stage is written to file.
          <VALUES>: 0

     -p, --print-similarity-measure-interval <unsignedIntegerValue>
          Prints out the CC similarity metric measure between the full-size input fixed
          and the transformed moving images at each iteration a value of 0 (the default)
          indicates that the full scale computation should not take placeany value greater
          than 0 represents the interval of full scale metric computation.
          <VALUES>: 0

     -v, --write-interval-volumes <unsignedIntegerValue>
          Writes out the output volume at each iteration. It helps to present the
          registration process as a short movie a value of 0 (the default) indicates that
          this option should not take placeany value greater than 0 represents the
          interval between the iterations which outputs are written to the disk.
          <VALUES>: 0

     -z, --collapse-output-transforms (1)/0
          Collapse output transforms. Specifically, enabling this option combines all
          adjacent transforms wherepossible. All adjacent linear transforms are written to
          disk in the forman itk affine transform (called xxxGenericAffine.mat).
          Similarly, all adjacent displacement field transforms are combined when written
          to disk (e.g. xxxWarp.nii.gz and xxxInverseWarp.nii.gz (if available)).Also, an
          output composite transform including the collapsed transforms is written to the
          disk (called outputCollapsed(Inverse)Composite).
          <VALUES>: 1

     -i, --initialize-transforms-per-stage (1)/0
          Initialize linear transforms from the previous stage. By enabling this option,
          the current linear stage transform is directly intialized from the previous
          stages linear transform; this allows multiple linear stages to be run where
          each stage directly updates the estimated linear transform from the previous
          stage. (e.g. Translation -> Rigid -> Affine).
          <VALUES>: 0

     -n, --interpolation Linear
                         NearestNeighbor
                         MultiLabel[<sigma=imageSpacing>,<alpha=4.0>]
                         Gaussian[<sigma=imageSpacing>,<alpha=1.0>]
                         BSpline[<order=3>]
                         CosineWindowedSinc
                         WelchWindowedSinc
                         HammingWindowedSinc
                         LanczosWindowedSinc
          Several interpolation options are available in ITK. These have all been made
          available. Currently the interpolator choice is only used to warp (and possibly
          inverse warp) the final output image(s).

     -g, --restrict-deformation PxQxR
          This option allows the user to restrict the optimization of the displacement
          field, translation, rigid or affine transform on a per-component basis. For
          example, if one wants to limit the deformation or rotation of 3-D volume to the
          first two dimensions, this is possible by specifying a weight vector of '1x1x0'
          for a deformation field or '1x1x0x1x1x0' for a rigid
          transformation.Low-dimensional restriction only works if there are no preceding
          transformations.

     -q, --initial-fixed-transform initialTransform
                                   [initialTransform,<useInverse>]
                                   [fixedImage,movingImage,initializationFeature]
          Specify the initial fixed transform(s) which get immediately incorporated into
          the composite transform. The order of the transforms is stack-esque in that the
          last transform specified on the command line is the first to be applied. In
          addition to initialization with ITK transforms, the user can perform an initial
          translation alignment by specifying the fixed and moving images and selecting an
          initialization feature. These features include using the geometric center of the
          images (=0), the image intensities (=1), or the origin of the images (=2).

     -r, --initial-moving-transform initialTransform
                                    [initialTransform,<useInverse>]
                                    [fixedImage,movingImage,initializationFeature]
          Specify the initial moving transform(s) which get immediately incorporated into
          the composite transform. The order of the transforms is stack-esque in that the
          last transform specified on the command line is the first to be applied. In
          addition to initialization with ITK transforms, the user can perform an initial
          translation alignment by specifying the fixed and moving images and selecting an
          initialization feature. These features include using the geometric center of the
          images (=0), the image intensities (=1), or the origin of the images (=2).

     -m, --metric CC[fixedImage,movingImage,metricWeight,radius,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  MI[fixedImage,movingImage,metricWeight,numberOfBins,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  Mattes[fixedImage,movingImage,metricWeight,numberOfBins,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  MeanSquares[fixedImage,movingImage,metricWeight,radius=NA,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  Demons[fixedImage,movingImage,metricWeight,radius=NA,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  GC[fixedImage,movingImage,metricWeight,radius=NA,<samplingStrategy={None,Regular,Random}>,<samplingPercentage=[0,1]>]
                  ICP[fixedPointSet,movingPointSet,metricWeight,<samplingPercentage=[0,1]>,<boundaryPointsOnly=0>]
                  PSE[fixedPointSet,movingPointSet,metricWeight,<samplingPercentage=[0,1]>,<boundaryPointsOnly=0>,<pointSetSigma=1>,<kNeighborhood=50>]
                  JHCT[fixedPointSet,movingPointSet,metricWeight,<samplingPercentage=[0,1]>,<boundaryPointsOnly=0>,<pointSetSigma=1>,<kNeighborhood=50>,<alpha=1.1>,<useAnisotropicCovariances=1>]
                  IGDM[fixedImage,movingImage,metricWeight,fixedMask,movingMask,<neighborhoodRadius=0x0>,<intensitySigma=0>,<distanceSigma=0>,<kNeighborhood=1>,<gradientSigma=1>]
          These image metrics are available--- CC: ANTS neighborhood cross correlation,
          MI: Mutual information, Demons: (Thirion), MeanSquares, and GC: Global
          Correlation. The "metricWeight" variable is used to modulate the per stage
          weighting of the metrics. The metrics can also employ a sampling strategy
          defined by a sampling percentage. The sampling strategy defaults to 'None' (aka
          a dense sampling of one sample per voxel), otherwise it defines a point set over
          which to optimize the metric. The point set can be on a regular lattice or a
          random lattice of points slightly perturbed to minimize aliasing artifacts.
          samplingPercentage defines the fraction of points to select from the domain. In
          addition, three point set metrics are available: Euclidean (ICP), Point-set
          expectation (PSE), and Jensen-Havrda-Charvet-Tsallis (JHCT).

     -t, --transform Rigid[gradientStep]
                     Affine[gradientStep]
                     CompositeAffine[gradientStep]
                     Similarity[gradientStep]
                     Translation[gradientStep]
                     BSpline[gradientStep,meshSizeAtBaseLevel]
                     GaussianDisplacementField[gradientStep,updateFieldVarianceInVoxelSpace,totalFieldVarianceInVoxelSpace]
                     BSplineDisplacementField[gradientStep,updateFieldMeshSizeAtBaseLevel,totalFieldMeshSizeAtBaseLevel,<splineOrder=3>]
                     TimeVaryingVelocityField[gradientStep,numberOfTimeIndices,updateFieldVarianceInVoxelSpace,updateFieldTimeVariance,totalFieldVarianceInVoxelSpace,totalFieldTimeVariance]
                     TimeVaryingBSplineVelocityField[gradientStep,velocityFieldMeshSize,<numberOfTimePointSamples=4>,<splineOrder=3>]
                     SyN[gradientStep,updateFieldVarianceInVoxelSpace,totalFieldVarianceInVoxelSpace]
                     BSplineSyN[gradientStep,updateFieldMeshSizeAtBaseLevel,totalFieldMeshSizeAtBaseLevel,<splineOrder=3>]
                     Exponential[gradientStep,updateFieldVarianceInVoxelSpace,velocityFieldVarianceInVoxelSpace,<numberOfIntegrationSteps>]
                     BSplineExponential[gradientStep,updateFieldMeshSizeAtBaseLevel,velocityFieldMeshSizeAtBaseLevel,<numberOfIntegrationSteps>,<splineOrder=3>]
          Several transform options are available. The gradientStep or learningRate
          characterizes the gradient descent optimization and is scaled appropriately for
          each transform using the shift scales estimator. Subsequent parameters are
          transform-specific and can be determined from the usage. For the B-spline
          transforms one can also specify the smoothing in terms of spline distance (i.e.
          knot spacing).

     -c, --convergence MxNxO
                       [MxNxO,<convergenceThreshold=1e-6>,<convergenceWindowSize=10>]
          Convergence is determined from the number of iterations per level and is
          determined by fitting a line to the normalized energy profile of the last N
          iterations (where N is specified by the window size) and determining the slope
          which is then compared with the convergence threshold.

     -s, --smoothing-sigmas MxNxO...
          Specify the sigma of gaussian smoothing at each level. Units are given in terms
          of voxels ('vox') or physical spacing ('mm'). Example usage is '4x2x1mm' and
          '4x2x1vox' where no units implies voxel spacing.

     -f, --shrink-factors MxNxO...
          Specify the shrink factor for the virtual domain (typically the fixed image) at
          each level.

     -u, --use-histogram-matching
          Histogram match the images before registration.

     -l, --use-estimate-learning-rate-once
          turn on the option that lets you estimate the learning rate step size only at
          the beginning of each level. * useful as a second stage of fine-scale
          registration.

     -w, --winsorize-image-intensities [lowerQuantile,upperQuantile]
          Winsorize data based on specified quantiles.

     -x, --masks [fixedImageMask,movingImageMask]
          Image masks to limit voxels considered by the metric.

     --float
          Use 'float' instead of 'double' for computations.
          <VALUES>: 0

     -v, --verbose (0)/1
          Verbose output.

     -h
          Print the help menu (short version).

     --help
          Print the help menu. Will also print values used on the current command line
          call.
          <VALUES>: 1
~~~~~~~~~~~~~~~~~~

## However,

[binaries are available](https://github.com/stnava/ANTs/releases).

![](./installation/figures/binaries.png)

## Steps for binary installation

1. Download appropriate binaries (~350 MB) from ANTs [website](https://github.com/stnava/ANTs/releases).

2. De-archive downloaded file to desired directory.

3. <a href="#/from-the-terminal">export</a> the ```ANTSPATH``` to the .profile or .bashrc file.

4. <a href="#/now-test-installation">Test</a> the installation by calling an ANTs program help menu.