### 2015-05-08 installation/antsr_installation.Rmd

# _ANTsR_ installation

## R

http://www.r-project.org

## Requirements for source installation

* [CMake](http://www.cmake.org) -- build system
* [git](http://git-scm.com/downloads) -- version control system
* C++ compiler

##

Install _ANTsR_ dependencies

~~~~~~~~~~~~~~~~~
$ R

R version 3.2.0 (2015-04-16) -- "Full of Ingredients"
Copyright (C) 2015 The R Foundation for Statistical Computing
Platform: x86_64-apple-darwin13.4.0 (64-bit)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type 'license()' or 'licence()' for distribution details.

  Natural language support but running in an English locale

R is a collaborative project with many contributors.
Type 'contributors()' for more information and
'citation()' on how to cite R or R packages in publications.

Type 'demo()' for some demos, 'help()' for on-line help, or
'help.start()' for an HTML browser interface to help.
Type 'q()' to quit R.

> antsrDependencies <- c( "Rcpp", "tools", "methods", "magrittr" )
> install.packages( pkgs = antsrDependencies, dependencies = TRUE )
~~~~~~~~~~~~~~~~~~

## Command line install

~~~~~~~~~~~~~~~~~
$ cd ${WHICH_PATH}
$ git clone https://github.com/stnava/ITKR.git
$ git clone https://github.com/stnava/ANTsR.git
~~~~~~~~~~~~~~~~~

## Install ITK and ANTsR

~~~~~~~~~~~~~~~~~
$ R CMD INSTALL ITKR
$ R CMD INSTALL ANTsR
~~~~~~~~~~~~~~~~~

## ANTsR updates?

~~~~~~~~~~~~~~~~~
$ cd ${WHICH_PATH}/ITKR
$ git pull
$ cd ${WHICH_PATH}/ANTsR
$ git pull
$ R CMD INSTALL ITKR
$ R CMD INSTALL ANTsR
~~~~~~~~~~~~~~~~~

## However,

[binaries are available](https://github.com/stnava/ANTsR/releases).

![](./installation/figures/antsrBinaries.png)

## Steps for binary installation

* Download [archive](https://github.com/stnava/ANTsR/releases) to the desired location
* install using R

~~~~~~~~~~~~~~~~~~~~~~~~~~
$ R CMD INSTALL ANTsR*gz
~~~~~~~~~~~~~~~~~~~~~~~~~~

## Test installation

~~~~~~~~~~~~~~~~~~~~~~~~~~
$ R
.
.
.
> library( ANTsR )
~~~~~~~~~~~~~~~~~~~~~~~~~~

_if no error message, then done!_

##

Let's do something simple:

1. Fetch MNI template.

2. Read in antsImage

3. Perform N3* bias correction

##

_Cross your fingers..._

~~~~~~~~~~~~~~~~~~~~~~~~~~
$ R
.
.
.
> library( ANTsR )
> mniTemplateFileName <- getANTsRData( "mni" )
trying URL 'http://placid.nlm.nih.gov/download?items=10785'
Content type 'application/octet-stream' length 3219212 bytes (3.1 MB)
==================================================
downloaded 3.1 MB
> image <- antsImageRead( mniTemplateFileName )
> imageN3 <- n3BiasFieldCorrection( image, 4 )
~~~~~~~~~~~~~~~~~~~~~~~~~~

## RStudio

(add more here)