### 2015-05-08 installation/snap_installation.Rmd

# ITK-SNAP installation

##

![](./installation/figures/snaplogo3.png)

## MacOS

1. [Download](http://www.itksnap.org/pmwiki/pmwiki.php?n=Downloads.SNAP3)

2. Drag the ```ITK-SNAP.app``` icon to the Applications folder

3. Drag the ```itksnap``` icon to the ```usr_local_bin``` folder

## Linux

1. [Download](http://www.itksnap.org/pmwiki/pmwiki.php?n=Downloads.SNAP3)

2. Open terminal.

3. Uncompress the downloaded file and copy the archive directory to ```/usr/local/```.

~~~~~~~~~~~~~~~~~~~~~~~
$ tar -zxvf itksnap-3.2.0-rc2-20140919-Linux_x86_64.tar.gz
$ sudo cp itksnap-3.2.0-rc2-20140919-Linux_x86_64/ /usr/local/
$ itksnap
~~~~~~~~~~~~~~~~~~~~~~~

## Thanks Paul!

and [others](http://www.itksnap.org/pmwiki/pmwiki.php?n=Main.Credits)