### 2015-05-09 evaluation/antsvfreesurfer2_extra.Rmd

# But, wait, there's more!

##

[__Data availability__](https://github.com/ntustison/KapowskiChronicles)

*  "_Hey, can I have the FreeSurfer measurements for the entorhinal cortex?_"  <span style="color:blue;">Sure, why not?</span>

*  "_I was wondering if you would make available the CT in each ROI for each subject..._" <span style="color:blue;">Sure, why not?  Would you also like the surface areas and volumes?</span>

*  "_Can I have one or more of the templates that you used for your study?_"  <span style="color:blue;">Would you like the priors as well?</span>

*  "_Hey Nick, can we hire you to do an [EZ-MAP]-type study?_" <span style="color:blue;">What about something better?</span>

##  What about using both the T1 & T2?

##

__ANTs tools are multivariate__

<div align="center"><img height="450" src="./evaluation/figures/rfRmse0.5ANTsT1T2.png" frameborder="0"></img></div>

##

__Arno-thick-in-the-head (ATITH)__

<br><br>
<span style="color:blue;">_What if we made a crude estimate of the cortical thickness?_</span>
<br><br>

$thickness_{ROI} = \frac{volume_{ROI}}{area_{ROI}}$

##

<span style="color:blue;">Make sure you estimate the [surface area](http://www.insight-journal.org/browse/publication/852) correctly!</span>

![](./evaluation/figures/LabelThickness.png)

## So how does ATITH compare with ANTs, FS?

##

__Prediction from cortical thickness data:<br>gender and age__

<div align="center"><img height="350" src="./evaluation/figures/atith.png" frameborder="0"></img></div>

##

![](./evaluation/figures/brainConstellationMaps.png)

# But the best part is ...

## it is absolutely free!

~~~~~~~~~~~~~~~~~~~~~~~
${ANTSPATH}/antsCorticalThickness.sh \
  -a IXI/T1/IXI002-Guys-0828-T1.nii.gz \
  -e IXI/template/T_template0.nii.gz \
  -m IXI/template/T_template0ProbabilityMask.nii.gz \
  -f IXI/template/T_template0ExtractionMask.nii.gz \
  -p IXI/template/Priors/priors%d.nii.gz \
  -o IXI/ANTsResults/IXI002-Guys-0828-
~~~~~~~~~~~~~~~~~~~~~~~

_(as in speech, not beer)_

## Don't worry,

we'll run a 2-D example this afternoon.
<BR><BR>

https://github.com/ntustison/antsCorticalThicknessExample

# Extensions