### 2015-05-09 lesions/lesions.Rmd

# Cortical thickness with lesions

##
```antsCorticalThickness.sh``` is adaptable!

1. Register subject (or single subject template) to normal template with priors.

2. Transform lesion mask to normal template.

3. Create additional "lesion" prior, i.e.

~~~~~~~~~~~~~~~~~~~~~~~
SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1
~~~~~~~~~~~~~~~~~~~~~~~

4.  Subtract out lesion prior from all other priors and keep values $\in [0,1]$.

##

__Modified template spatial priors__

<div align="center"><img height="450" src="lesions/figures/lesionPrior.png" frameborder="0"></img></div>

##

```antsCorticalThickness.sh``` using lesion prior

<div align="center"><img height="450" src="lesions/figures/lesionSegmentation.png" frameborder="0"></img></div>

<!--
Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."
-->