### 2015-05-09 longitudinal/longitudinal2.Rmd

# Longitudinal processing

##

![](./longitudinal/figures/unbiasedPapers.png)

##

![](./longitudinal/figures/longitudinalPipeline.png)