### 2015-05-17 background/background.Rmd
ENH:  More material.

# ANTs lineage

##

__Image mapping & perception: 1877-8__
<BR><BR>
Francis Galton: _Can we see criminality in the face?_

![](./figures/galton.png)

<!--
Notes:
* Use of "composite portraiture" and what he called "eugenics"
* Can we see criminality, mental illness, syphilis, and others in facial features?
* Despite the obvious prejudices driving his research, he was a pioneer.
-->

##

<!--
![](./background/figures/FrancisGalton2.jpg)
-->

<!--
Notes:
* Another example where he has manually categorized facial features and try to combine them
* Did not find much---composite tends to blend away to normality.
-->

##

Speaking of criminality...<BR>
_Can we say anything about the U.S. Congress?_

![](./background/figures/congressPersons.png)
__Maybe he should have used [ANTs](https://github.com/ntustison/CongressionalFaceTemplates)?__

<!--
Notes:
-->

##

__Image mapping & biology: 1917__
<BR><BR>
D''Arcy Thompson: _Comparison of related forms_

![](./figures/dthompson.png)

<!--
Notes:
* Book: On Growth and Form
* Descriptive not susceptible to hypothesis testing
* "This book of mine has little need of preface, for indeed it is 'all preface' from beginning to end."
-->

##

__Image mapping & biology: Current__
<BR><BR>
_Look where we are now!_
<BR><BR>
![](./background/figures/r16xr85.png)

<!--
Notes:
* Book: On Growth and Form
* Descriptive not susceptible to hypothesis testing
* "This book of mine has little need of preface, for indeed it is 'all preface' from beginning to end."
-->

##

__ANTs family tree__

![](./figures/lineage.jpg)

<!--
References: @Horn1981, @Gee1993, @Grenander1993, @Thompson2001, @Miller2002, @Shen2002, @Arnold2014, @Thirion1998, @Rueckert1999, @Fischl2012, @Ashburner2012
-->

##

__Initial scope__

![](./figures/ants_initial_scope.png)

<!--
Notes:
* Brian work with Murray Grossman
* First registration paper
  - WBIR submission
  - FEM registraiton using B-spline basis functions.
  - "yeah, that registration is not that good---Jim would not like it."
-->

##

__Diffeomorphisms__:plausible physical modeling

<div align="center"><img src="./figures/sillyputty.png" frameborder="0"></img></div>

"differentiable map with differentiable inverse"

##

__Fine-grained and flexible maps__

![](./figures/highresdiffeos.jpg)

<!--
... to correct a misconception about diffeomorphisms ...
-->

##

 __Diffeomorphisms__: image parameterization in a metric space

![](./figures/shape_appearance.png)

##

__General purpose library for multivariate image registration, segmentation & statistical analysis tools__

* 170,000+ lines of C++, 6$+$ years of work, 15+ collaborators.

* Generic mathematical methods that are tunable for
application specific domains.

* <span style="color:red;">no free lunch</span>: <span style="color:black;">_An algorithm must use prior knowledge about a problem to do well on that problem._</span>

<!--
* Deep testing on multiple platforms ... osx, linux, windows.

* Several "wins" in public knock-abouts ( [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496), [Murphy 2011](http://www.ncbi.nlm.nih.gov/pubmed/21632295), [SATA 2012 and 2013](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837555/), [BRATS 2013](http://martinos.org/qtim/miccai2013/proc_brats_2013.pdf), others )
-->

## Still quite a bit to do

##

__On documentation__

![](figures/Chainsaw.jpg)

<!--
Notes:
* ANTs documentation is only slightly better than what is shown here.
-->

##

Developers can be blind to doc deficiencies while users are blind to what we provide!

![](figures/blindmen.jpg)