### 2015-05-17 background/people.Rmd

# Founding developers

##

![](./figures/brian_and_edie_sedgwick.jpg)

<!--
Notes:
* Where is this from?
* Brain--visionary,brilliant.
* New college in Florida --- degree in Physics and Literature(?)
* Went to Siemens and got kicked out, fortunately, by ???  "You need to get a doctorate."
* Jim Gee first graduate student
* Jim wanted to work on FEM registration
  - side tangent:  Jim''s notebooks of FEM registration code
-->

##

![](./background/figures/brucelee.png)

<!--
NOtes:
* Not much to say
  - did graduate work at the University of Virginia (hyper. gas lungs)
  - Washington University in St. Louis (cardiac)
  - actually bombed interview with Siemens
  - 2004: Jim offered a job without an interview (long term collaboration)
  - wrote graph cuts for ITK in a month
-->

##
<BR><BR>

<div align="center"><img width="1800" src="./figures/brian_and_nick.png" frameborder="0"></img></div>

<!--
NOtes:
* Worked together for almost 10 years
* have learned tons
* have written quite a bit of code together
-->

##

![](./figures/antscollab.jpg)

$+$ [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

<!--
NOtes:
* However, it is not just us
* Other people have done quite a bit
* and others who are not shown such as Arman
-->