### 2015-05-17 evaluation/antsvfreesurfer2_extra.Rmd

# But, wait, there's more!

##

[__Data availability__](https://github.com/ntustison/KapowskiChronicles)

*  "_Hey, can I have the FreeSurfer measurements for the entorhinal cortex?_"  <span style="color:blue;">Sure, why not?</span>

*  "_I was wondering if you would make available the CT in each ROI for each subject..._" <span style="color:blue;">Sure, why not?  Would you also like the surface areas and volumes?</span>

*  "_Can I have one or more of the templates that you used for your study?_"  <span style="color:blue;">Would you like the priors as well?</span>

*  "_Hey Nick, can we hire you to do an [EZ-MAP]-type study?_" <span style="color:blue;">What about something better?</span>

##  What about using both the T1 & T2?

##

__ANTs tools are multivariate__

<div align="center"><img height="450" src="./evaluation/figures/rfRmse0.5ANTsT1T2.png" frameborder="0"></img></div>

##

__Arno-thick-in-the-head (ATITH)__

<br><br>
<span style="color:blue;">_What if we made a crude estimate of the cortical thickness?_</span>
<br><br>

$thickness_{ROI} = \frac{volume_{ROI}}{area_{ROI}}$

##

<span style="color:blue;">Make sure you estimate the [surface area](http://www.insight-journal.org/browse/publication/852) correctly!</span>

![](./evaluation/figures/LabelThickness.png)

## So how does ATITH compare with ANTs, FS?

##

__Prediction from cortical thickness data:<br>gender and age__

<div align="center"><img height="350" src="./evaluation/figures/atith.png" frameborder="0"></img></div>

##

![](./evaluation/figures/brainConstellationMaps.png)

## And currently brewing on the UVa cluster...

_Does [denoising](http://www.ncbi.nlm.nih.gov/pubmed/20027588) help?_

##

* ``itkAdaptiveNonLocalMeansDenoisingImageFilter.h``

* ``itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx``

~~~~~~~~~~~~~~~~~~~~~~
$ DenoiseImage

COMMAND:
     DenoiseImage
          Denoise an image using a spatially adaptive filter originally described in J. V.
          Manjon, P. Coupe, Luis Marti-Bonmati, D. L. Collins, and M. Robles. Adaptive
          Non-Local Means Denoising of MR Images With Spatially Varying Noise Levels,
          Journal of Magnetic Resonance Imaging, 31:192-203, June 2010.

OPTIONS:
     -d, --image-dimensionality 2/3/4
          This option forces the image to be treated as a specified-dimensional image. If
          not specified, the program tries to infer the dimensionality from the input
          image.

     -i, --input-image inputImageFilename
          A scalar image is expected as input for noise correction.

     -n, --noise-model (Rician)/Gaussian
          Employ a Rician or Gaussian noise model.

     -s, --shrink-factor (1)/2/3/...
          Running noise correction on large images can be time consuming. To lessen
          computation time, the input image can be resampled. The shrink factor, specified
          as a single integer, describes this resampling. Shrink factor = 1 is the
          default.

     -o, --output correctedImage
                  [correctedImage,<noiseImage>]
          The output consists of the noise corrected version of the input image.
          Optionally, one can also output the estimated noise image.

     --version
          Get Version Information.

     -v, --verbose (0)/1
          Verbose output.

     -h
          Print the help menu (short version).

     --help
          Print the help menu.
~~~~~~~~~~~~~~~~~~~~~~

# But the best part is ...

## it is absolutely free!

~~~~~~~~~~~~~~~~~~~~~~~
${ANTSPATH}/antsCorticalThickness.sh \
  -a IXI/T1/IXI002-Guys-0828-T1.nii.gz \
  -e IXI/template/T_template0.nii.gz \
  -m IXI/template/T_template0ProbabilityMask.nii.gz \
  -f IXI/template/T_template0ExtractionMask.nii.gz \
  -p IXI/template/Priors/priors%d.nii.gz \
  -o IXI/ANTsResults/IXI002-Guys-0828-
~~~~~~~~~~~~~~~~~~~~~~~

_(as in speech, not beer)_

## Don't worry,

we'll run a 2-D example this afternoon.
<BR><BR>

https://github.com/ntustison/antsCorticalThicknessExample

# Extensions