### 2015-05-17 longitudinal/longitudinal2.Rmd

# Longitudinal processing

##

![](./longitudinal/figures/unbiasedPapers.png)

##

![](./longitudinal/figures/longitudinalPipeline.png)

##

``antsLongitudinalCorticalThickness.sh``

~~~~~~~~~~~~~~~~~~~~
$ antsLongitudinalCorticalThickness.sh

antsLongitudinalCorticalThickness.sh performs a longitudinal cortical thickness estimation.  The following steps
are performed:
  1. Create a single-subject template (SST) from all the data
  2. Create priors for the SST
     a. Run the SST through the individual cortical thickness pipeline.
     b. The brain extraction SST prior is created by smoothing the brain extraction
        mask created during 2a.
     c. If labeled atlases are not provided, we smooth the posteriors from 2a to create
        the SST segmentation priors, otherwise we use antsMalfLabeling to create a set of
        posteriors (https://github.com/ntustison/antsCookTemplatePriorsExample).
  3. Using the SST + priors, we run each subject through the antsCorticalThickness
     pipeline.

Usage:

antsLongitudinalCorticalThickness.sh -d imageDimension
              -e brainTemplate
              -m brainExtractionProbabilityMask
              -p brainSegmentationPriors
              <OPTARGS>
              -o outputPrefix
              ${anatomicalImages[@]}

Example:

  bash /Users/ntustison/Pkg/ANTs/bin/bin/antsLongitudinalCorticalThickness.sh -d 3 -e brainWithSkullTemplate.nii.gz -m brainPrior.nii.gz -p segmentationPriors%d.nii.gz -o output ${anatomicalImages[@]}

Required arguments:

     -d:  Image dimension                       2 or 3 (for 2- or 3-dimensional image)
     -e:  Brain template                        Anatomical *intensity* template (possibly created using a population
                                                data set with buildtemplateparallel.sh in ANTs).  This template is
                                                *not* skull-stripped.
     -m:  Brain extraction probability mask     Brain *probability* mask created using e.g. LPBA40 labels which
                                                have brain masks defined, and warped to anatomical template and
                                                averaged resulting in a probability image.
     -p:  Brain segmentation priors             Tissue *probability* priors corresponding to the image specified
                                                with the -e option.  Specified using c-style formatting, e.g.
                                                -p labelsPriors%02d.nii.gz.  We assume that the first four priors
                                                are ordered as follows
                                                  1:  csf
                                                  2:  cortical gm
                                                  3:  wm
                                                  4:  deep gm
     -o:  Output prefix                         The following subdirectory and images are created for the single
                                                subject template
                                                  * ${OUTPUT_PREFIX}SingleSubjectTemplate/
                                                  * ${OUTPUT_PREFIX}SingleSubjectTemplate/T_template*.nii.gz

     anatomical images                          Set of multimodal input data assumed to be specified ordered as
                                                follows:
                                                  ${time1_modality1} ${time1_modality2} ... ${time1_modalityN} \
                                                  ${time2_modality1} ${time2_modality2} ...
                                                  .
                                                  .
                                                  .
                                                  ${timeN_modality1} ...

						A single modality is expected by default, in which case the input images
						are simply ordered by time:

						  ${time1_modality1} ${time2_modality1} ... ${timeN_modality1}

						If there are multiple modalities, use the -k option to specify how many.

Optional arguments:

     -s:  image file suffix                     Any of the standard ITK IO formats e.g. nrrd, nii.gz (default), mhd
     -c:  control type                          Control for parallel computation (default 0):
                                                  0 = run serially
                                                  1 = SGE qsub
                                                  2 = use PEXEC (localhost)
                                                  3 = Apple XGrid
                                                  4 = PBS qsub
     -t:  template for t1 registration          Anatomical *intensity* template (assumed to be skull-stripped).  A common
                                                use case would be where this would be the same template as specified in the
                                                -e option which is not skull stripped.
     -a:                                        Atlases (assumed to be skull-stripped) used to cook template priors.  If atlases
                                                aren't used then we simply smooth the single-subject template posteriors after
                                                passing through antsCorticalThickness.sh. Example:

 						  -a atlas1.nii.gz -a atlas2.nii.gz ... -a atlasN.nii.gz

     -l:                                        Labels associated with each atlas, in the same order as they are specified
						with the -a option. The number of labels in each image is assumed to be equal
                                                to the number of priors.
     -f:  extraction registration mask          Mask (defined in the template space) used during registration
                                                for brain extraction.
     -j:  number of cpu cores                   Number of cpu cores to use locally for pexec option (default 2; requires "-c 2")
     -k:  number of modalities                  Number of modalities used to construct the template (default 1):  For example,
                                                if one wanted to use multiple modalities consisting of T1, T2, and FA
                                                components ("-k 3").
     -n:  use SST cortical thickness prior      If set to '1', the cortical thickness map from the single-subject template is used
                                                as a prior constraint for each of the individual calls to antsCorticalThickness.sh
                                                (default = 0).
     -g:  use floating-point precision          Use floating point precision in registrations (default = 0)
     -v:  Atropos segmentation weight (SST)     Atropos spatial prior *probability* weight for the segmentation for the single
                                                subject template (default = 0.25)
     -w:  Atropos segmentation weight (Indiv.)  Atropos spatial prior *probability* weight for the segmentation for the individual
                                                time points (default = 0.5)
     -q:  Use quick registration parameters     If = 1, use antsRegistrationSyNQuick.sh as the basis for registration
                                                during brain extraction, brain segmentation, and (optional) normalization
                                                to a template.  Otherwise use antsRegistrationSyN.sh (default = 0).
     -r:  rigid alignment to SST                This option dictates if the individual subjects are registered to the single
                                                subject template before running through antsCorticalThickness.  This potentially
                                                reduces bias caused by subject orientation and voxel spacing (default = 0).
     -z:  Test / debug mode                     If > 0, runs a faster version of the script. Only for testing. Implies -u 0.
                                                Requires single thread computation for complete reproducibility.
~~~~~~~~~~~~~~~~~~~~