### 2015-05-17 retrospective/retrospective.Rmd

---
title: <span style="color:red;">ANTs:</span><br />A retrospective
author: Brian (<span style="color:blue;">PENN</span>) and <br />Nick (<span style="color:orange;">UVA</span>)
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

```{r global options, include=FALSE}
library(rmarkdown)
library(knitr)
figpath<-'figures/'
opts_chunk$set(fig.width=9, fig.height=6, fig.path=figpath,
               warning=FALSE, message=FALSE)
```

<!--
Notes:
* Acknowledgments
* Honor to be here
-->

This talk is online at [http://stnava.github.io/ANTs2015/](http://stnava.github.io/ANTs2015/) with colored [links](http://stnava.github.io/ANTs2015/) meant to be clicked for more information.

<!--
Notes:
* Format:  we chose this particular format as it gives us quite a bit of flexibility
  in reusability, public availability, and linking to additional online information
  - Apologize if it doesn't conform to your sense of aesthetics.  We empathize but...
-->