### 2015-05-20 background/background.Rmd
ENH:  Adding mni tutorial stuff.

# ANTs lineage

## Image mapping and perception: 1877

Francis Galton: _Can we see criminality in the face?_

![](./figures/galton.png)

_What about syphilis, mental illness?_

<!--
Notes:
* Use of "composite portraiture" and what he called "eugenics"
* Can we see criminality, mental illness, syphilis, and others in facial features?
* Despite the obvious prejudices driving his research, he was a pioneer.
-->

<!--

##

![](./background/figures/FrancisGalton2.jpg)
-->

<!--
Notes:
* Another example where he has manually categorized facial features and try to combine them
* Did not find much---composite tends to blend away to normality.
-->

##  Speaking of criminality...

_Can we say anything about the U.S. Congress?_

![](./background/figures/congressPersons.png)

__Maybe he should have used [ANTs](https://github.com/ntustison/CongressionalFaceTemplates)?__

<!--
Notes:
-->

##  Image mapping & biology:  1917

D'Arcy Thompson: _Comparison of related forms_

![](./figures/dthompson.png)

<!--
Notes:
* Book: On Growth and Form
* Descriptive not susceptible to hypothesis testing
* "This book of mine has little need of preface, for indeed it is 'all preface' from beginning to end."
-->

## Image mapping & biology: Current
<BR><BR>
_Look where we are now!_
<BR><BR>
![](./background/figures/r16xr85.png)

<!--
Notes:
* Book: On Growth and Form
* Descriptive not susceptible to hypothesis testing
* "This book of mine has little need of preface, for indeed it is 'all preface' from beginning to end."
-->

## ANTs family tree

![](./figures/lineage.jpg)

<!--
References: @Horn1981, @Gee1993, @Grenander1993, @Thompson2001, @Miller2002, @Shen2002, @Arnold2014, @Thirion1998, @Rueckert1999, @Fischl2012, @Ashburner2012
-->

## Initial scope

![](./figures/ants_initial_scope.png)

<!--
Notes:
* Brian work with Murray Grossman
* First registration paper
  - WBIR submission
  - FEM registraiton using B-spline basis functions.
  - "yeah, that registration is not that good---Jim would not like it."
-->

## Diffeomorphisms:  Occam's razor modeling

![](./figures/sillyputty.png)

_differentiable map with differentiable inverse_

## Diffeomoprhisms:  fine-grained and flexible maps

![](./figures/highresdiffeos.jpg)

<!--
... to correct a misconception about diffeomorphisms ...
-->

## Diffeomorphisms: image parameterization in a metric space

![](./figures/shape_appearance.png)