### 2015-05-20 retrospective/neuroimage_fmri.Rmd

# Retrospective

##

## ANTs evolution

![](./retrospective/figures/retrospective.jpg)

<!--
Notes:
* No equations, just kind of a look back at where ANTs came from.
* Manuscripts are a bit deceptive in that they often present a linear development
  whereas science is messy (e.g. Popper, Carnap, and others).  Review from Fred Bookstein
  said the same thing---more on that later.
-->

## FreeSurfer, Bruce Fischl

![](./retrospective/figures/juicyDetails.png)

<!--
Notes:
* NeuroImage 26(2)  "20 years of fmri"
* FreeSurfer, Bruce Fischl "only manuscript where the reviewer requested more juicy details of the history of the development of a set of ideas."
* Yearly course with FSL, 2003 in Santa Monica "Where's the free surfer course, Dude?"
* As a quasi-relevant tangent, not a surfer, decade long playful argument with significant other
-->

<!--

## Other juicy bits

![](./retrospective/figures/logos.png)

Notes:
* Additional interesting bits of information
* Logos
* Why a fish skeleton?
* "FOSSIL vs EFF ESS ELL"
* What is the ice cream cone for and what does it have to do with British humour?
-->

## NeuroImaging research as literature

"Contingency.
AFNI is one result of unexpected events.
I met a woman and fell in love."

<!--
Notes:
* Love this!
* Perhaps not in the realm of "It was the best of times it was the worst of times" or "Call me Ishmael."
but, in our field, that's as close as we're going to get!
* His love interest was faculty at the University of Milwaukee.
* I certainly do not expect these stories or discussions to be retained in memory
but I hope they point to the fact that Brian and I love what we do, we love reproducible
science and enjoy helping other people, and that ANTs is a set of tools that might be
worth investigating.
-->

## ANTs--_begotten in SyN_

![](./retrospective/figures/Welcometotherealm.jpg)

<!--
Notes:
* In somewhat playful antithetical terms, ANTs was beggotten in SyN.
-->