### 2015-05-20 evaluation/antsvfreesurfer2_extra.Rmd

# But, wait, there's more!

##

[__Data availability__](https://github.com/ntustison/KapowskiChronicles)

*  "_Hey, can I have the FreeSurfer measurements for the entorhinal cortex?_"  Sure, why not?

*  "_I was wondering if you would make available the CT in each ROI for each subject..._" Sure, why not?  Would you also like the surface areas and volumes?

*  "_Can I have one or more of the templates that you used for your study?_" Would you like the priors as well?

##  What about using both the T1 & T2?

## ANTs tools are multivariate

![](./evaluation/figures/rfRmse05ANTsT1T2.png)

## Arno-thick-in-the-head (ATITH)

_What if we made a crude estimate of the cortical thickness?_

$$thickness_{ROI} = \frac{volume_{ROI}}{area_{ROI}}$$

## Make sure you estimate the [surface area](http://www.insight-journal.org/browse/publication/852) correctly!

![](./evaluation/figures/LabelThickness.png)

## So how does ATITH compare with ANTs, FS?

## Prediction from cortical thickness data

![](./evaluation/figures/atith.png)

## Brain constellation maps

![](./evaluation/figures/brainConstellationMaps.png)

## And currently brewing on the UVa cluster...

_Does [denoising](http://www.ncbi.nlm.nih.gov/pubmed/20027588) help?_

## Contribution from Jose

* ``itkAdaptiveNonLocalMeansDenoisingImageFilter.h``

* ``itkAdaptiveNonLocalMeansDenoisingImageFilter.hxx``

```bash
$ DenoiseImage

COMMAND:
     DenoiseImage
          Denoise an image using a spatially adaptive filter originally described in J. V.
          Manjon, P. Coupe, Luis Marti-Bonmati, D. L. Collins, and M. Robles. Adaptive
          Non-Local Means Denoising of MR Images With Spatially Varying Noise Levels,
          Journal of Magnetic Resonance Imaging, 31:192-203, June 2010.

OPTIONS:
     -d, --image-dimensionality 2/3/4
          This option forces the image to be treated as a specified-dimensional image. If
          not specified, the program tries to infer the dimensionality from the input
          image.

     -i, --input-image inputImageFilename
          A scalar image is expected as input for noise correction.

     -n, --noise-model (Rician)/Gaussian
          Employ a Rician or Gaussian noise model.

     -s, --shrink-factor (1)/2/3/...
          Running noise correction on large images can be time consuming. To lessen
          computation time, the input image can be resampled. The shrink factor, specified
          as a single integer, describes this resampling. Shrink factor = 1 is the
          default.

     -o, --output correctedImage
                  [correctedImage,<noiseImage>]
          The output consists of the noise corrected version of the input image.
          Optionally, one can also output the estimated noise image.

     --version
          Get Version Information.

     -v, --verbose (0)/1
          Verbose output.

     -h
          Print the help menu (short version).

     --help
          Print the help menu.
```

# But the best part is ...

## it is absolutely free!

```bash
${ANTSPATH}/antsCorticalThickness.sh \
  -a IXI/T1/IXI002-Guys-0828-T1.nii.gz \
  -e IXI/template/T_template0.nii.gz \
  -m IXI/template/T_template0ProbabilityMask.nii.gz \
  -f IXI/template/T_template0ExtractionMask.nii.gz \
  -p IXI/template/Priors/priors%d.nii.gz \
  -o IXI/ANTsResults/IXI002-Guys-0828-
```

_(as in speech, not beer)_

## We'll discuss more this afternoon

https://github.com/ntustison/antsCorticalThicknessExample