### 2015-05-20 lesions/lesions.Rmd

# Cortical thickness with lesions

## ```antsCorticalThickness.sh``` is adaptable!

1. Register subject (or single subject template) to normal template with priors.

2. Transform lesion mask to normal template.

3. Create additional "lesion" prior, i.e. ``SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1``

4.  Subtract out lesion prior from all other priors and keep values $\in [0,1]$.

## Modified template spatial priors

![](lesions/figures/lesionPrior.png)

## ```antsCorticalThickness.sh``` using lesion prior

![](lesions/figures/lesionSegmentation.png)

<!--
Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."
-->