### 2015-05-20 mystyle.tex

\pgfdeclareimage[height=0.1cm, width=1cm]{logo}{}
\logo{\pgfuseimage{logo}}
\institute{}
\definecolor{links}{HTML}{2A1B81}
\definecolor{mypink2}{RGB}{219, 48, 122}
\hypersetup{colorlinks,linkcolor=links,urlcolor=mypink2}
\usefonttheme{professionalfonts}
\setbeamerfont{note page}{family*=pplx,size=\footnotesize} % Palatino for notes
\usepackage{helvet}
\useinnertheme{rectangles}