### 2015-05-20 summary/summary.Rmd

# ANTs:  in summary

## Show me the code!

__General purpose library for multivariate image registration, segmentation & statistical analysis tools__

* 170,000+ lines of C++, 6$+$ years of work, 15+ collaborators.

* Generic mathematical methods that are tunable for
application specific domains.

* <span style="color:red;">no free lunch</span>: <span style="color:black;">_An algorithm must use prior knowledge about a problem to do well on that problem._</span>

<!--
* Deep testing on multiple platforms ... osx, linux, windows.

* Several "wins" in public knock-abouts ( [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496), [Murphy 2011](http://www.ncbi.nlm.nih.gov/pubmed/21632295), [SATA 2012 and 2013](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837555/), [BRATS 2013](http://martinos.org/qtim/miccai2013/proc_brats_2013.pdf), others )
-->

## Still quite a bit to do

## On documentation

![](figures/Chainsaw.jpg)

<!--
Notes:
* ANTs documentation is only slightly better than what is shown here.
-->

## We are blind to deficiencies.  Users blind to features.

![](figures/blindmen.jpg)