### 2016-07-17 competitions/competitions.Rmd

<!--
# Competitions
-->

<!--

## Algorithmic fight club

![](./competitions/figures/fightclub.jpg)

-->

## International competitions

* [Klein 2009](http://www.ncbi.nlm.nih.gov/pubmed/19195496): MRI brain registration

* [EMPIRE 2010](http://empire10.isi.uu.nl):  CT lung registration

* [Multi-Atlas Label Challenge 2012](https://masi.vuse.vanderbilt.edu/workshop2012/index.php/Main_Page):  MRI brain registration and segmentation

* [SATA Challenge 2013](https://masi.vuse.vanderbilt.edu/workshop2013/index.php/MICCAI_2013_SATA_Challenge_and_Workshop:Current_events): MRI cardiac and canine hind leg registration

* [BRATS 2013](http://martinos.org/qtim/miccai2013/): Multi-modal MRI brain segmentation

* [STACOM 2014 MoCo Challenge](http://www.cardiacatlas.org/web/stacom2014/moco-introduction):  MRI cardiac motion estimation

<!--- ![](./evaluation/figures/evalhistory.png) -->

<!--

## Klein, NeuroImage 2009

![](./competitions/figures/nihms-130679-f0001.jpg)

* 14 algorithms
* publicly available data
* different labeling protocols
* developers tune their own algorithms

## Klein results

![](./competitions/figures/KleinResults2009.png)

## Take-away message

_"One of the most significant findings of this study is that the relative performances of
the registration methods under comparison appear to be little affected by the choice of
subject population, labeling protocol, and type of overlap measure....  ART, SyN, IRTK, and
SPM's DARTEL Toolbox gave the best results according to overlap and distance measures,
with ART and SyN delivering the most consistently high accuracy across subjects and label sets."_

## EMPIRE 2010

![](./competitions/figures/EMPIRE10criteria.png)

* Alignment of lung boundaries,

* major fissures,

* annotated landmark pairs, and

* topology of displacement field.

## EMPIRE 2010 results

![](./competitions/figures/EMPIRE10Results.png)

## Multi-atlas 2012

![](./competitions/figures/NKI30_MALF.png)

## Multi-atlas 2012 results

![](./competitions/figures/MALF2012.png)

## SATA 2013

Three very different problem domains

![](./competitions/figures/SATA_data.png)

## BRATS 2013

![](./competitions/figures/brats2013results1.png)

## BRATS 2013 results

![](./competitions/figures/brats2013results2.png)

[Tustison, et al., Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with _ANTsR_, _Neuroinformatics_.](http://www.ncbi.nlm.nih.gov/pubmed/25433513)

## STACOM 2014

![](./competitions/figures/image_gallery.jpg)

-->

<!--- A great opportunity to showcase a [new SyN variant](http://www.ncbi.nlm.nih.gov/pubmed/24409140) -->