### 2016-07-17 licenses/licenses.Rmd

# Why would you care?

<!--

## \`\`Free beer\'\'

![](./licenses/figures/freebeerfreespeech.jpg)

-->

## Software for medical image analysis

* [FSL](http://fsl.fmrib.ox.ac.uk/fsldownloads/)

* [SPM](http://www.fil.ion.ucl.ac.uk/spm/software/)

* [FreeSurfer](http://www.fil.ion.ucl.ac.uk/spm/software/)

* [MIPAV](http://mipav.cit.nih.gov)

* [AFNI](https://afni.nimh.nih.gov/afni)

* Slicer, Elastix, SimpleITK, ANTs $\longleftrightarrow$ [Insight Toolkit](http://www.itk.org)

* Many more at [idoimaging.com](http://idoimaging.com)

<!--

## [FSL](http://fsl.fmrib.ox.ac.uk/fsldownloads/)

> No part of the Software may be reproduced, modified, transmitted or
> transferred in any form or by any means, electronic or mechanical,
> without the express permission of the University. The permission of
> the University is not required if the said reproduction, modification,
> transmission or transference is done without financial return, ...
> You may be held legally responsible for any copyright infringement
> that is caused or encouraged by your failure to abide by these terms
> and conditions.
>
> __You are not permitted under this Licence to use this Software
> commercially.__

## [SPM](http://www.fil.ion.ucl.ac.uk/spm/software/)

> SPM is free but copyright software, distributed under the terms of
> the GNU General Public Licence as published by the Free Software
> Foundation (either version 2, as given in file spm_LICENCE.man,
> or at your option, any later version). Further details on "copyleft"
> can be found at http://www.gnu.org/copyleft/. In particular, SPM is
> supplied as is. No formal support or maintenance is provided or implied.

> "Copyleft" says that anyone who redistributes the software, with or
> without changes, must pass along the freedom to further copy and change
> it. Copyleft guarantees that every user has freedom... __it is illegal to
> distribute the improved version except as free software__...

## [FreeSurfer](http://www.fil.ion.ucl.ac.uk/spm/software/)

> The license granted in this Software License includes without limitation
> the right to (i) incorporate the Software into proprietary programs
> (subject to any restrictions applicable to such programs), (ii) add
> your own copyright statement to your modifications of the Software, and
> (iii) provide additional or different license terms and conditions in your
> sublicenses of modifications of the Software; provided that in each case your
> use, reproduction or distribution of such modifications otherwise complies
> with the conditions stated in this Software License.

## [ANTs](https://github.com/stnava/ANTs/blob/master/ANTSCopyright.txt)

> Redistribution and use in source and binary forms, with or without
> modification, are permitted provided that the following conditions
> are met:
> 1. _Redistributions of source code must retain the above copyright
>   notice, this list of conditions and the following disclaimer._
> 2. _Redistributions in binary form must reproduce the above copyright
>   notice, this list of conditions and the following disclaimer in the
>   documentation and/or other materials provided with the distribution._
> 3. _Neither the name of the consortium nor the names of its contributors
>  may be used to endorse or promote products derived from this software
>  without specific prior written permission._

## [ITK](https://itk.org/ITK/project/licenseversion1.html)

> The license of the Insight Toolkit is an open-source Apache 2.0 license. It
> allows unrestricted use, including use in commercial products.
> Starting with ITK 4.0, The Insight Toolkit is distributed under the Apache 2.0
> license approved by the Open Source Initiative (OSI).

-->