### 2016-07-17 simpleItkTalk/itkv4.Rmd

# ITKv3 $\longrightarrow$ ITKv4

## Extensions of ITKv3

* [Elastix](http://elastix.isi.uu.nl)
* [Diffeomorphic Demons](http://www.insight-journal.org/browse/publication/154)
* [BrainsFit](http://www.insight-journal.org/browse/publication/180)
* [Advanced Normalization Tools (ANTs)](http://www.insight-journal.org/browse/publication/681)

## Basic components

* Optimizers (v4)
* Metrics (v4)
* Transforms
* Image registration methods (v4)

## Optimizers (v4) and scales estimators

* Scales estimator

## Metrics (v4)

* Intensity metrics
    * mean squares
    * mutual information (Mattes and joint histogram)
    * neighborhood cross-correlation
    * demons
    * global correlation
* Point-set metrics
    * iterative closest point
    * point-set expectation
    * Jensen-Havrda-Charvat-Tsallis
* Multi-metrics

## Transforms

* Linear (translation, rigid, \ldots)
* B-spline
* Displacement field (Gaussian, B-spline, SyN)
* Velocity field (exponential, time-varying)
* Composite

## Image registration methods (v4)

* $\rightarrow$ Input:
    * Fixed image and initial fixed transform
    * Moving image and initial moving transform
    * metric(s)
* $\leftarrow$ Output:
    * Optimized moving transform
* Special methods for SyN and velocity field transforms

## Beyond original SyN

\small

![](./papers/figures/Frontiers_ITK.png)

![](./papers/figures/Frontiers_BSplineSyN.png)