### 2017-04-24 antsr/antsr.Rmd
ENH:  ASNR 2017

# Advanced Normalization Tools in R (ANTsR)

## Multimodal Brain Tumor Segmentation (BRATS 2013)

\begin{centering}

\includegraphics[width=4.5in]{./competitions/figures/brats2013results1.png}

\end{centering}

\small

[Tustison, et al., Optimal symmetric multimodal templates and concatenated random forests for
supervised brain tumor segmentation (simplified) with _ANTsR_, _Neuroinformatics_.](http://www.ncbi.nlm.nih.gov/pubmed/25433513)

## White matter hyperintensities in TBI

\begin{centering}

\includegraphics[width=4.5in]{./wmhs/figures/wmhPipeline.png}

\end{centering}

## Social behavior and immunity dysfunction in mice

\begin{centering}

\includegraphics[width=3in]{./antsr/figures/filiano_rsfmri.png}

\end{centering}

## Other ANTsR work

* [Pediatric template of brain perfusion](http://www.nature.com/articles/sdata20153)

* [Automated segmentation of chronic stroke lesions using LINDA: Lesion identification with neighborhood data analysis](http://www.ncbi.nlm.nih.gov/pubmed/26756101)

* [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/25448483)

* Corrective learning for segmentation refinement