### 2017-04-24 background/background3.Rmd

# What is ANTs?

## NLM's The Visible Human Project and The Insight Toolkit

\begin{center}

\includegraphics[width=4.75in]{./figures/VisibleHumanProject_ITK.pdf}

\end{center}

## ANTs _began with a clinical research need._

\begin{center}

\includegraphics[width=4.75in]{./figures/ANTsEvolutionarySummary.pdf}

\end{center}

## Advanced Normalization Tools

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsNeuro.png}

\end{center}

## Advanced Normalization Tools $\rightarrow$ ``ITK-Lung''

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsLung.png}

\end{center}