### 2017-04-24 evaluation/antsvfreesurfer2.Rmd

# Putting it all together---the ANTs cortical thickness pipeline

## Cortical thickness studies

\scriptsize

```{r, results='asis', echo = FALSE}
conditions <- c( "Tetris-playing ability", "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                  "traumatic brain injury", "untreated male-to-female transsexuality" )

conditionsDF <- data.frame( Column1 = conditions[1:floor( 0.5*length(conditions) )], Column2 = conditions[( 1+floor( 0.5*length(conditions) ) ):length(conditions)] )
knitr::kable( conditionsDF )
```

<!--

## ANTs-related background

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
    + used other tools such as FMRIB Automated Segmentation Tool (FAST)
    + `KellySlater.cxx`
* pipedream scripts "adequate for Penn data."
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

## Basic components of the pipeline

1. template building (offline)
2. brain extraction
3. cortical thickness estimation
4. cortical parcellation

## Sample results

![](./evaluation/figures/components.png)

-->

## The ANTs structural brain mapping workflow

\begin{centering}

\includegraphics[width=4.5in]{./evaluation/figures/pipeline.png}

\end{centering}

## Template building

*Tailor data to your specific cohort*

\begin{centering}

\includegraphics[width=3in]{./evaluation/figures/templates.png}

\end{centering}

\small

* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.  
* Acquisition and anonymization (e.g. defacing) protocols are often different.

## Template priors

\begin{centering}

\includegraphics[width=2.5in]{./evaluation/figures/templatePriors.png}

\end{centering}

\small

Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.

## Brain extraction comparison

\begin{centering}

\includegraphics[width=3.5in]{./evaluation/figures/brainExtraction.png}

\end{centering}

\small

Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)

## Brain segmentation

\begin{centering}

\includegraphics[width=3.5in]{./evaluation/figures/brainSegmentation.png}

\end{centering}

\small

Randomly selected healthy individuals.  ```Atropos``` gets good performance across ages.

## Cortical thickness maps

\begin{centering}

\includegraphics[width=4in]{./evaluation/figures/corticalThicknessEstimation.png}

\end{centering}

\small

In contrast to FreeSurfer which warps coupled surface meshes to segment the gray matter,
*ANTs* diffeomorphically registers the white matter to the combined gray/white matters
while simultaneously estimating thickness.

----

\begin{centering}

\Large

{\bf But without ground truth, how does one evaluate the pipeline?}

\end{centering}

## Predict age and gender

\begin{centering}

$AGE \sim VOLUME + GENDER + \sum_{i=1}^{62} T(DKT_i)$

\end{centering}

<!--

## Open science principles

![](./evaluation/figures/ageDistribution.png)

* Public data sets (IXI, NKI, OASIS, MMRR)
* $>$ 1200 subjects, age 4 to over 94 years old

-->

## Prediction from cortical thickness data

\begin{centering}

\includegraphics[width=4in]{./evaluation/figures/evaluation.png}

\end{centering}

<!--

## Age prediction per site

\begin{centering}

\includegraphics[width=4in]{./evaluation/figures/agePredictionPerSite.png}

\end{centering}

-->

## Regional importance comparison

\begin{centering}

\small
ANTs (left) vs. FreeSurfer (right)

\includegraphics[width=4in]{./evaluation/figures/antsvfreesurfer_Importance.png}

\end{centering}

<!--

## Regional measurements

\begin{centering}

\includegraphics[width=4in]{./evaluation/figures/antsvfreesurfer_regionalPlots.png}

\end{centering}

-->