### 2018-02-15 background/background3.Rmd
ENH:  Add Cohen presentation.

# What is ANTs?

## NLM's The Visible Human Project and The Insight Toolkit

\begin{center}

\includegraphics[width=4.75in]{./figures/VisibleHumanProject_ITK.pdf}

\end{center}

## ANTs _began with a clinical research need._

\begin{center}

\includegraphics[width=4.75in]{./figures/ANTsEvolutionarySummary.pdf}

\end{center}

## Founding developers

\begin{center}
\includegraphics[width=4.75in]{./figures/brian_and_nick.png}
\end{center}

\begin{center}
BBA \& NJT
\end{center}

## Collaborators

\begin{center}

\includegraphics[width=4.75in]{./figures/antsCollaborators2.png}

\end{center}

$+$ [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

## Advanced Normalization Tools

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsNeuro.png}

\end{center}

## Advanced Normalization Tools $\rightarrow$ ``ITK-Lung''

\begin{center}

\includegraphics[width=4.75in]{./tools/figures/coreANtsToolsLung.png}

\end{center}